---
title: "`r config$publication_name`"
pagetitle: PCA England
always_allow_html: true
output:
  word_document:
    toc: yes
    toc_depth: '2'
  html_document:
    anchor_sections: no
    css: www/style.css
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: no
---

<script>
  $(document).ready(function() {
    $('#TOC').prepend('<div><img src=`r knitr::image_uri("www/national-statistics-logo.svg")` alt=\"National Statistics logo\" style=\"position: relative; width: 50%; margin: 10% auto 10% auto; display: block;\"></div>');
  });
  
  $(document).ready(function() {
    $('#header').prepend('<img src=`r knitr::image_uri("www/bsa_logo.svg")` alt=\"NHSBSA logo\" style=\"position:absolute; top: 0; right: 0; padding: 10px; width: 25%;\">');
  });
</script>

<html lang="en">

```{r setup, include=FALSE}

# set code chunk options to disable echo by default
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE)


```

<main>

# `r config$publication_sub_name` {.toc-ignore}

`r paste("Published", config$publication_date)`

::::{.row style="display: flex; padding-bottom: 10px;"} 
:::{.col-md-12} 
`r infoBox_border(header = "Changes to these statistics", text = paste0("CHANGES HERE. We are interested in any feedback about the publication or these changes, which you can send by using our ", htmltools::HTML(paste("<a href='",config$stats_survey_link,"'>Official Statistics feedback survey</a>")),"."), width = "100%")` 
::: 
::::

## Key findings

::::{.row style="display: flex; padding-bottom: 10px;"} 
:::{.col-md-6} 
`r infoBox_no_border(header = "", text = "Prescription items dispensed in the community in England cost £10.4 billion.", width = "100%")` 
::: 
:::{.col-md-6} 
`r infoBox_no_border(header = "", text = "1.18 billion items were dispensed in the community in England.", width = "100%")` 
:::
::::

The cost of prescription items dispensed in the community in England was £10.4 billion, an 8% increase from £9.69 billion in 2021/22.

The number of prescription items dispensed in the community in England was 1.18 billion, a 3% increase from 1.14 billion in 2021/22.

The most dispensed drug in England in 2022/23 was Atorvastatin, with 59 million items, while Apixaban was the drug with the highest costs of £430 million.

Apixaban 5mg tablets had the largest increase in cost between 2021/22 and 2022/23 in England, with an absolute increase of £280 million.

---

## 1. Introduction

### 1.1. Scope {.toc-ignore}

Prescription Cost Analysis (PCA) includes information about the costs and volumes of prescription items that have been dispensed in England.

These statistics cover prescription items that were prescribed in England, Scotland, Wales, Northern Ireland, and the Channel Islands that were subsequently dispensed in the community in England and submitted to the NHS Business Services Authority (NHSBSA) for reimbursement. When a prescription is 'dispensed in the community', it means that it was dispensed at a pharmacy, by and applicance contractor, dispensing doctor, or as a personally administered item. However, if a prescription item has been prescribed in England but dispensed elsewhere in the United Kingdom, it has not been included in these statistics.

This publication is a National Statistic release. National Statistics status means that PCA meets the highest standards of trustworthiness, quality, and public value and complies with all aspects of the [Code of Practice for Statistics](https://code.statisticsauthority.gov.uk/).

The designation of this publication as a National Statistic was confirmed in May 2021 following an assessment by the Office for Statistics Regulation (OSR). You can read more about this confirmation and the assessment of these statistics on the [OSR website](https://osr.statisticsauthority.gov.uk/correspondence/ed-humpherson-to-matthew-wilson-confirmation-of-national-statistics-designation-for-prescription-cost-analysis-england-statistics/).

### 1.2. Definitions {.toc-ignore}

::::{.row style="display: flex; padding-bottom: 10px;"} 
:::{.col-md-12} 
`r infoBox_border(header = "Item", text = "A single unit of medication listed separately on a prescription form. In this publication, an example of an item would be Paracetamol 500mg x28.", width = "100%")` 
::: 
::::

::::{.row style="display: flex; padding-bottom: 10px;"} 
:::{.col-md-12} 
`r infoBox_border("NIC", text = "The Net Ingredient Cost (NIC) is the basic price of the medication and the quantity prescribed. It does not include other fees incurred by dispensing contractors, such as controlled drug fees or the single activity fee. The basic price is determined by the [Drug Tariff](https://www.nhsbsa.nhs.uk/pharmacies-gp-practices-and-appliance-contractors/drug-tariff) or by the manufacturer, wholesaler, or supplier of the product.", width = "100%")`
::: 
::::

::::{.row style="display: flex; padding-bottom: 10px;"} 
:::{.col-md-12} 
`r infoBox_border("Classification", text = "This publication uses the British National Formulary (BNF), which lists medicines used in the UK and classifies them according to their primary therapeutic use. Medication may also be prescribed for other purposes beyond their intended use. For example, certain antidepressants may be recommended to individuals experiencing chronic pain. The NHSBSA does not capture the clinical indication of a prescription during processing.", width = "100%")`
::: 
::::

::::{.row style="display: flex; padding-bottom: 10px;"} 
:::{.col-md-12} 
`r infoBox_border("Generic prescribing", text = "This is when a drug is prescribed using its non-proprietary name instead of a specific brand name. This allows for any suitable drug that meets the same standards as the brand name drug to be dispensed, which can result in cost savings.", width = "100%")`
::: 
::::

::::{.row style="display: flex; padding-bottom: 10px;"} 
:::{.col-md-12} 
`r infoBox_border("Branded prescribing", text = "This is when a drug is prescribed using a specific brand name.", width = "100%")`
::: 
::::

---

## 2. Results and commentary

### `r paste0("Figure 1: Cost of prescription items dispensed in England 2014/15 to ", paste0(substr(max_data_fy,1,5), substr(max_data_fy, 8, 9)))`

```{r total_nic, out.width = "100%"}

figure_1

```
`r get_download_button(title = "Download chart data", data = figure_1_data, filename = "figure_1")`

Source: [Summary Tables - Prescription Cost Analysis - National](`r config$fy_nat_link`)

::::{.row style="display: flex; padding-bottom: 10px;"} 
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "Prescription items dispensed in the community in England cost £10.4 billion.", width = "100%")` 
::: 
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "There was an 8% increase in costs from 2021/22 to 2022/23.", width = "100%")` 
:::
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "2022/23 is the fourth consecutive year that costs have increased.", width = "100%")` 
:::
::::

In England, the cost of prescription items in 2022/23 was £10.4 billion, which is 8% more than the previous year's cost of £9.69 billion. Compared to 2014/15, the cost has increased by 17% or £1.48 billion. This marks the fourth year in a row where the cost of items dispensed has increased, after seeing a decrease in the previous three years from 2015/16 to 2018/19.

### `r paste0("Figure 2: Number of prescription items dispensed in England 2014/15 to ", paste0(substr(max_data_fy,1,5), substr(max_data_fy, 8, 9)))`

```{r total_items, out.width = "100%"}

figure_2

```
`r get_download_button(title = "Download chart data", data = figure_2_data, filename = "figure_2")`

Source: [Summary Tables - Prescription Cost Analysis - National](`r config$fy_nat_link`)

::::{.row style="display: flex; padding-bottom: 10px;"} 
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "1.18 billion items were dispensed in the community in England.", width = "100%")` 
::: 
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "There was a 3% increase in items from 2021/22 to 2022/23.", width = "100%")` 
:::
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "Items dispensed continue to trend upwards.", width = "100%")` 
:::
::::

In England, there were 1.18 billion items dispensed in 2022/23, which is 3% more than the previous year's 1.14 billion items. Compared to 2014/15, the number of items has increased by 10%.

### `r paste0("Figure 3: Distribution of items dispensed in England amongst BNF chapters ", paste0(substr(max_data_fy,1,5), substr(max_data_fy, 8, 9)))`

```{r bnf_chapters_items, out.width = "100%"}

figure_3

```
`r get_download_button(title = "Download chart data", data = figure_3_data, filename = "figure_3")`

Source: [Summary Tables - Prescription Cost Analysis - Chapter](`r config$fy_nat_link`)


::::{.row style="display: flex; padding-bottom: 10px;"} 
:::{.col-md-12} 
`r infoBox_border(header = "", text = "These statistics use a system called the British National Formulary (BNF) to group medicines togther based on their primary theraputic indication. The BNF has multiple levels, starting with the largest grouping called chapters, follwed by sections, paragraphs, sub-paragraphs, chemical substances, products, and individual presentations.<br><br>The NHSBSA uses the BNF classification system implemented prior to the release of edition 70, which includes 6 additional chapters (18 to 23) created by NHS Prescription Services. These chapters are used to classify products that fall outside the chapters 1 to 15, and mainly include dressing, appliances and medical devices.<br><br>However, presentations in chapters 20 to 23 do not have assigned BNF paragraphs, sub-paragraphs, chemical substances, or products.", width = "100%")` 
::: 
::::

::::{.row style="display: flex; padding-bottom: 10px;"} 
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "BNF chapter 02 - Cardiovascular System accounted for 29% of all items dispensed.", width = "100%")` 
::: 
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "BNF chapter 04 - Central Nervous System accounted for 19% of all items dispensed.", width = "100%")` 
:::
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "BNF chapter 06 - Endocrine System accounted for 11% of all items dispensed.", width = "100%")` 
:::
::::

In 2022/23, the chapter on the Cardiovascular System in the BNF had the highest number of prescription items dispensed in England, with a total of 340 million items, accounting for 29% of all items dispensed. The Central Nevous System chapter followed with 230 million items, representing 19% of items. The Endocrine System chapter had 130 million items, accounting for 11% of items.

### `r paste0("Figure 4: Distribution of cost of items dispensed in England amongst BNF chapters ", paste0(substr(max_data_fy,1,5), substr(max_data_fy, 8, 9)))`

```{r bnf_chapters_costs, out.width = "100%"}

figure_4

```
`r get_download_button(title = "Download chart data", data = figure_4_data, filename = "figure_4")`

Source: [Summary Tables - Prescription Cost Analysis - Chapter](`r config$fy_nat_link`)

::::{.row style="display: flex; padding-bottom: 10px;"} 
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "BNF chapter 06 - Endocrine System accounted for 17% of total costs.", width = "100%")` 
::: 
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "BNF chapter 02 - Cardiovascular System accounted for 16% of total costs.", width = "100%")`
:::
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "BNF chapter 04 - Central Nervous System accounted for 15% of total costs.", width = "100%")` 
:::
::::

In 2022/23, the BNF chapter on the Endocrine System had the highest costs, totaling £1.78 billion, which was equal to 17% of all costs. The Cardiovascular System chapter followed with costs of £1.67 billion, representing 16% of costs. The Central Nervous System chapter had costs of £1.59 billion, 15% of the total costs.

Despite having 63% fewer items, the Endocrine System had costs that were 6% higher than the Cardiovascular System during 2022/23.

### `r paste0("Figure 5: Proportion of items prescribed generically and proportion of costs excluding appliances England ", paste0(substr(max_data_fy,1,5), substr(max_data_fy, 8, 9)))`

```{r generic_presc, out.width = "100%"}

figure_5

```
`r get_download_button(title = "Download chart data", data = figure_5_data, filename = "figure_5")`

Source: [Additional Tables - Prescription Cost Analysis - table A5](`r config$add_anl_link`)

::::{.row style="display: flex; padding-bottom: 10px;"} 
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "85% of prescription items prescibed generically in 2022/23.", width = "100%")` 
::: 
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "Rate of generic prescrbing has remained consistent between 2014/15 and 2022/23.", width = "100%")` 
:::
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "Generic prescribing accounted for 62% of costs in 2022/23.", width = "100%")` 
:::
::::

In 2022/23, 85% of prescription items were prescribed generically, which is slightly higher then the previous year. However, the rate of generic prescribing has remained consistent over the period. There has been a 1 percentage point variation in the period between 2014/15 and 2022/23. Note that these figures exclude preparation class 4 products - appliances.

Regarding the cost of reimbursement, in 2022/23, generic prescribing accounted for 62% of the total cost. This is a increase of 1 percentage point from the previous year. The proportion of costs attributed to generic prescribing has shown greater variation, with a range of 7 percentage points in the period 2014/15 to 2022/23. 

### `r paste0("Figure 6: Sankey chart to show the flow of prescribing and dispensing volumes England ", paste0(substr(max_data_fy,1,5), substr(max_data_fy, 8, 9)))`

```{r presc_sankey, out.width = "100%", out.height=400}

figure_6

```
`r get_download_button(title = "Download chart data", data = figure_6_data, filename = "figure_6")`

### `r paste0("Figure 7: The 10 most dispensed chemical substances by total items in ", paste0(substr(max_data_fy,1,5), substr(max_data_fy, 8, 9)))`

```{r top_10_items, out.width = "100%"}

figure_7

```
`r get_download_button(title = "Download chart data", data = figure_7_data, filename = "figure_7")`

Source: [Additional Tables - Prescription Cost Analysis - A3](`r config$add_anl_link`)

::::{.row style="display: flex; padding-bottom: 10px;"} 
:::{.col-md-12} 
`r infoBox_border(header = "", text = "A chemical substance is the registered name for the active ingredient in a medicine, such as Omeparzole. Different presentations of the same active ingredient are grouped together under the chemical substance name. For instance, Omeprazole 20mg and Omeprazole 40mg capsules both contain the same active ingredient. Products like dressing and appliances in BNF chapters 20 to 23 do not have a chemical substance assigned to them", width = "100%")` 
::: 
::::

::::{.row style="display: flex; padding-bottom: 10px;"} 
:::{.col-md-6} 
`r infoBox_no_border(header = "", text = "Atorvastatin was the most dispensed chemical susbtance with 59 million items.", width = "100%")` 
::: 
:::{.col-md-6} 
`r infoBox_no_border(header = "", text = "Four of the top ten chemical substances by items belong to BNF chapter 02 - Cardiovascular System.", width = "100%")` 
:::
::::

[Atorvastatin](https://www.nhs.uk/medicines/atorvastatin/) was the most dispensed chemical substance in 2022/23, with 59 million items. This medication is a type of statin that is commonly used to lower cholesterol and help prevent heart disease. Four of the top ten dispensed chemical substances belong to BNF chapter 02 - Cardiovascular System. These substances are Atorvastatin, Amlodipine, Ramipril, and Bisoprolol fumarate.

### `r paste0("Figure 8: The 10 most dispensed chemical substances by cost in ", paste0(substr(max_data_fy,1,5), substr(max_data_fy, 8, 9)))`

```{r top_10_costs, out.width = "100%"}

figure_8

```
`r get_download_button(title = "Download chart data", data = figure_8_data, filename = "figure_8")`

Source: [Additional Tables - Prescription Cost Analysis - A2](`r config$add_anl_link`)

::::{.row style="display: flex; padding-bottom: 10px;"} 
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "Apixiban had the highest costs of £430 million.", width = "100%")` 
::: 
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "Three of the top ten chemical substances by costs are anticoagulants that arte used to help prevent blood clots.", width = "100%")` 
:::
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "Two of the top ten chemical substances by costs are used to treat repiratory illnesses.", width = "100%")` 
:::
::::

Apixiban had the highest cost for items dispensed in the community in England, with a total cost of £400 million. [Apixaban](https://www.nhs.uk/medicines/apixaban/) is an anticoagulant that is primarily used to treat individuals who have had a health problem caused by a blood clot or to prevent future blood clots.

The top ten chemical substances by cost included three drugs that are commonly used to help prevent blood clots. These are Apixaban, Rivaroxaban, and Edoxaban.

The top ten chemical substances by cost included two drugs that are commonly used to treat respiratory illnesses. These are Beclometasone dipropionate and Budesonide.

### `r paste0("Figure ",9,": Dispensed items per capita by Sustainability and Transformation Partnership (ICB) ",year_tidy)`

```{r items_per_pop, out.width = "100%"}

figure_9
    
```
`r get_download_button(title = "Download chart data", data = figure_9_data, filename = "figure_9")`

Source: [Summary Tables - Prescription Cost Analysis - ICB National](`r config$fy_icb_link`) | [Office for National Statistics - Mid-year population estimates](https://www.ons.gov.uk/peoplepopulationandcommunity/populationandmigration/populationestimates/datasets/clinicalcommissioninggroupmidyearpopulationestimates)

::::{.row style="display: flex; padding-bottom: 10px;"} 
:::{.col-md-12} 
`r infoBox_border(header = "", text = paste("Integrated care boards (ICB) are statutory NHS organisations which are responsible for developing a plan for meeting the health needs of the population, managing the NHS budget and arranging for the provision of health services in a geographical area. <br><br>However, it's important to note that the ICB listed in these statistics refers to the dispensing contractor as listed in the NHSBSA admistrative records, rather than the prescribing organisation of an item.<br><br>Prescribing and dispensing trends can vary regionally, and there are many factors that can influence these decisions. Some of the factors that can affect prescribing and dispensing trends include population demographics, prevalence of chronic disease, and the presence of specialist dispensing contractors within a given area.<br><br>If you're interested in learning more about ICBs and how they were used in these statistics, please refer to the [Background Information and Methodology note](",config$background_link,") that accompanies this release."), width = "100%")` 
::: 
::::

::::{.row style="display: flex; padding-bottom: 10px;"} 
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "The mean number of items dispensed per capita in England was 21 in 2022/23.", width = "100%")` 
::: 
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "NHS West Yorkshire Integrated Care Board had the highest dispensed items per capita with 29.", width = "100%")` 
:::
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "NHS Surrey Heartlands Integrated Care Board had the lowest with 13.", width = "100%")` 
:::
::::

### `r paste0("Figure ",10,": Cost per capita (£) by Sustainability and Transformation Partnership (ICB) ",year_tidy)`

```{r nic_per_pop, out.width = "100%"}

figure_10

```
`r get_download_button(title = "Download chart data", data = figure_10_data, filename = "figure_10")`

Source: [Summary Tables - Prescription Cost Analysis - ICB National](`r config$fy_icb_link`) | [Office for National Statistics - Mid-year population estimates](https://www.ons.gov.uk/peoplepopulationandcommunity/populationandmigration/populationestimates/datasets/clinicalcommissioninggroupmidyearpopulationestimates)

::::{.row style="display: flex; padding-bottom: 10px;"} 
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "The mean cost of items dispensed per capita in England was £184 in 2022/23.", width = "100%")` 
::: 
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "NHS Cambridgeshire and Peterborough Integrated Care Board had the highest cost per capita of £390.", width = "100%")` 
:::
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "NHS South East London Integrated Care BoardP had the lowest of £123.", width = "100%")` 
:::
::::

It is worth noting that NHS Cambridgeshire and Peterborough Integrated Care Board offer specialist stoma care, 34% of their total costs of £340 million in 2022/23 were in BNF chapter 23 (Stoma Appliances).

### `r paste0("Figure 11: Top 10 presentations by percentage increase in unit cost ",paste0(substr(max_data_fy_minus_1,1,5), substr(max_data_fy_minus_1, 8, 9)), " to ", paste0(substr(max_data_fy,1,5), substr(max_data_fy, 8, 9)))`

```{r unit_cost_increase, out.width = "100%"}

figure_11

```
`r get_download_button(title = "Download chart data", data = figure_11_data, filename = "figure_11")`


Source: [Additional Tables - Prescription Cost Analysis - A11](`r config$add_anl_link`)

::::{.row style="display: flex; padding-bottom: 10px;"} 
:::{.col-md-12} 
`r infoBox_border(header = "", text = "The chart in this section has been limited to presentations with a total cost in 2022/23 greater than £1 million.<br><br>A unit cost for a presentation is calculated by dividing the total cost of items dispensed in the period by the total quantity.", width = "100%")` 
::: 
::::

::::{.row style="display: flex; padding-bottom: 10px;"} 
:::{.col-md-12} 
`r infoBox_no_border(header = "", text = "Chloral hydrate 500mg/5ml oral solution/mixture BP 2000 increased in unit cost the most in 2022/23 by 2,024%.", width = "100%")` 
::: 
::::

### `r paste0("Figure 12: Top 10 presentations by percentage decrease in unit cost ",paste0(substr(max_data_fy_minus_1,1,5), substr(max_data_fy_minus_1, 8, 9)), " to ", paste0(substr(max_data_fy,1,5), substr(max_data_fy, 8, 9)))`

```{r unit_cost_decrease, out.width = "100%"}

figure_12

```
`r get_download_button(title = "Download chart data", data = figure_12_data, filename = "figure_12")`


Source: [Additional Tables - Prescription Cost Analysis - A12](`r config$add_anl_link`)

::::{.row style="display: flex; padding-bottom: 10px;"} 
:::{.col-md-12} 
`r infoBox_border(header = "", text = "The chart in this section has been limited to presentations with a total cost in 2022/23 greater than £1 million.<br><br>A unit cost for a presentation is calculated by dividing the total cost of items dispensed in the period by the total quantity.", width = "100%")` 
::: 
::::

::::{.row style="display: flex; padding-bottom: 10px;"} 
:::{.col-md-12} 
`r infoBox_no_border(header = "", text = "Salofalk 1g/application foam enema decreased in unit cost the most in 2022/23 by 78%.", width = "100%")` 
::: 
::::


### `r paste0("Figure 13: Top 10 presentations by increase in total cost ", paste0(substr(max_data_fy_minus_1,1,5), substr(max_data_fy_minus_1, 8, 9)), " to ", paste0(substr(max_data_fy,1,5), substr(max_data_fy, 8, 9)))`

```{r abs_cost_increase, out.width = "100%"}

figure_13

```
`r get_download_button(title = "Download chart data", data = figure_13_data, filename = "figure_13")`

Source: [Additional Tables - Prescription Cost Analysis - A13](`r config$add_anl_link`)

::::{.row style="display: flex; padding-bottom: 10px;"} 
:::{.col-md-12} 
`r infoBox_border(header = "", text = "The chart in this section has been limited to presentations with a total cost in 2022/23 greater than £1 million.", width = "100%")` 
::: 
::::


::::{.row style="display: flex; padding-bottom: 10px;"} 
:::{.col-md-12} 
`r infoBox_no_border(header = "", text = "Between 2021/22 and 2022/23 the total costs of Apixaban 5mg tablets increased by £280 million, the highest absolute change of all presentations. LINE HERE ABOUT WHEN APIXIBAN BECAME GENERIC - JUST CHECK FOR EXACT DATE ON THIS.", width = "100%")` 
::: 
::::

### `r paste0("Figure 14: Top 10 presentations by decrease in total cost ", paste0(substr(max_data_fy_minus_1,1,5), substr(max_data_fy_minus_1, 8, 9)), " to ", paste0(substr(max_data_fy,1,5), substr(max_data_fy, 8, 9)))`

```{r abs_cost_decrease, out.width = "100%"}

figure_14

```
`r get_download_button(title = "Download chart data", data = figure_14_data, filename = "figure_14")`

Source: [Additional Tables - Prescription Cost Analysis - A14](`r config$add_anl_link`)

::::{.row style="display: flex; padding-bottom: 10px;"} 
:::{.col-md-12} 
`r infoBox_border(header = "", text = "The chart in this section has been limited to presentations with a total cost in 2021/22 greater than £1 million.", width = "100%")` 
::: 
::::

::::{.row style="display: flex; padding-bottom: 10px;"} 
:::{.col-md-12} 
`r infoBox_no_border(header = "", text = "Between 2021/22 and 2022/23 the total costs of Eliquis 5mg tablets decreased by £260 million, the highest absolute change of all presentations. AS ABOVE LINE ABOUT GENERIC APIXIBAN AS ELQUIS IS BRANDED VERSION.", width = "100%")` 
::: 
::::

---

## 3. UK comparisons

::::{.row style="display: flex; padding-bottom: 10px;"} 
:::{.col-md-12} 
`r infoBox_border(header = "", text = paste("Prescription Cost Analysis statistics are produced seperately by each UK nation, but there are differences in data collection, healthcare structures, and methodolgy. Therefore, it's important to be cautious when comparing PCA data between different UK nations and to consider the broader context of the UK's health and social care landscape.<br><br>We've selected come measures that enable broad comparisons between the different deveolved authorities, but for more detailed information, please refer to each country's individual publication.<br><br>We based our comparisons on the latest available PCA data and mid-year population estimates from the Office for National Statistics (ONS), for the financail year 2021/22, using 2020 mid-year population estimates to calculate per capita measures. For Wales, the data relates to prescriptions issued by Welsh GPs and dispensed in Wales or other UK countries, as we did not have figures for prescriptions dispensed solely in Wales at the time of publication.<br><br>For more information on the PCA publications of each UK nation and the differences between them, please read the [Background Information and Methodology note](",config$background_link,") released alongside this publication."), width = "100%")` 
::: 
::::

### `r paste0("Figure 15: Items per capita for the UK nations ", max_data_fy_minus_1)`

```{r uk_comparisons_items, out.width = "100%"}

figure_15

```
`r get_download_button(title = "Download chart data", data = figure_15_data, filename = "figure_15")`

Source: [Additional Tables - Prescription Cost Analysis - A1](`r config$add_anl_link`) | [Public Health Scotland - Dispenser payments and prescription cost analysis](`r config$scotland_pca`) | [Welsh Government - Prescriptions in Wales](`r config$wa_pca_link`) | [HSC Business Services Organisation - General pharmaceutical services and prescribing statistics](`r config$ni_pca`) | [Office for National Statistics - Mid-year population estimates](https://www.ons.gov.uk/peoplepopulationandcommunity/populationandmigration/populationestimates/datasets/populationestimatestimeseriesdataset)

::::{.row style="display: flex; padding-bottom: 10px;"} 
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "20 items per head of population were dispensed in England in 2021/22.", width = "100%")` 
::: 
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "This was 1 item more than Scotland who had the fewest items dispensed per capita of 19.", width = "100%")` 
::: 
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "26 items per head of population were dispensed in Wales in 2021/22.", width = "100%")` 
::: 
::::

### `r paste0("Figure 16: Cost per capita for the UK nations ", max_data_fy_minus_1)`

```{r uk_comparisons_costs, out.width = "100%"}

figure_16


```
`r get_download_button(title = "Download chart data", data = figure_16_data, filename = "figure_16")`

Source: [Additional Tables - Prescription Cost Analysis - A1](`r config$add_anl_link`) | [Public Health Scotland - Dispenser payments and prescription cost analysis](`r config$scotland_pca`) | [Welsh Government - Prescriptions in Wales](`r config$wa_pca_link`) | [HSC Business Services Organisation - General pharmaceutical services and prescribing statistics](`r config$ni_pca`) | [Office for National Statistics - Mid-year population estimates](https://www.ons.gov.uk/peoplepopulationandcommunity/populationandmigration/populationestimates/datasets/populationestimatestimeseriesdataset)

::::{.row style="display: flex; padding-bottom: 10px;"} 
:::{.col-md-6} 
`r infoBox_no_border(header = "", text = "England had the lowest prescription item cost per head of population in 2020/21 of £171.", width = "100%")` 
::: 
:::{.col-md-6} 
`r infoBox_no_border(header = "", text = "This was £76 less than Northern Ireland at £247 per capita.", width = "100%")` 
::: 
::::

---

## 4. About these statistics

### 4.1. Monthly PCA data {.toc-ignore}

The NHSBSA publishes a [monthly administrative management information dataset](`r config$nhsbsa_pca_link`) for PCA using the same methodology as these statistics however, it is not an Official Statistic.

There's a slight difference between the annual National Statistics and the monthly administrative data for PCA. The annual publication uses the geographical structure as of the 31st March for that year, while the monthly data uses the structure at the end of the month. If a dispensing contractor changes parent organisation during the year, this may affect the ICB level total in the aggregated monthly data compared to the annual National Statistics, but it won't affect the national totals. However, such cases are rare.


### 4.2. Uses of these statistics {.toc-ignore}

These statistics serve several purposes, including:

* providing a national overview of community prescription costs and volumes in England for the calendar year

* analysing cost and volume trends across different levels of the BNF hierarchy

* allowing public scrutiny of national prescribing habits

* monitoring uptake of new to market proprietary presentations

* monitoring the proportion of prescriptions dispensed generically

### 4.3. Changes to this publication {.toc-ignore}

For full details of the changes made please refer to the [Background Information and Methodology note](`r config$background_link`) released alongside these statistics.

### 4.4. Timeliness of this publication {.toc-ignore}

This publication has historically been released in the month of April. To accommodate the inclusion of financial year information these statistics will now be released in June, after March data becomes available each year. Calendar year data is currently released at the same time as financial year information. We are going to get user’s views on publishing calendar year data earlier.

---

## 5. Rounding

The high-level figures in this statistical summary have been rounded as per the table below:

```{r rounding_table, out.width = "100%"}
rounding_table <- data.table(
  "From" = c("0", "1,001", "10,001", "100,001", "1,000,001", "10,000,001", "100,000,001"),
  "To" = c("1,000", "10,000", "100,000", "1,000,000", "10,000,000", "100,000,000", "100,000,000,000"),
  "Round to nearest" = c("1", "100", "1,000", "10,000", "100,000", "1,000,000", "10,000,000")
)

knitr::kable(rounding_table, align = "rrr")%>%
  kable_styling(bootstrap_options = c("striped"))
```

All changes and totals are calculated prior to rounding. Percentage changes are calculated prior to rounding and then are rounded to the nearest whole number. As all figures within this statistical summary have been rounded, they may not add up.

The summary tables released with this publication allow users to investigate this data at lower levels of granularity. Figures in the supplementary tables have not been rounded.

---

## 6. Accessibility

### 6.1. How to request content in an accessible format {.toc-ignore}

If you need information on this website in a different format like accessible PDF, large print, easy read, audio recording or braille, you can contact us by:

**Email**: nhsbsa.accessibility@nhs.net

**Phone**: 0191 203 5318

[Find out about call charges](https://www.nhsbsa.nhs.uk/contact-us/call-charges-and-phone-numbers)

We’ll consider your request and get back to you in 5 working days.

These contact details are only for accessibility queries. This inbox is not for technical queries or IT problems. If you have a query that is not about accessibility, go to the ‘Contact us’ section of this page.

---

## 7. Feedback and contact us

Feedback is important to us. We welcome any questions and comments relating to these statistics. 

You can complete a [short survey about this publication](`r config$stats_survey_link`) to help us improve the Official Statistics that we produce. All responses will remain anonymous, and individuals will not be identifiable in any report that we produce.

You can view our [privacy policy](https://www.nhsbsa.nhs.uk/our-policies/privacy) on our website to see how your data is used and stored.

You can contact us by:

**Email:** nhsbsa.statistics@nhs.net

**You can also write to us at:**

NHSBSA - Statistics  
NHS Business Services Authority  
Stella House  
Goldcrest Way  
Newburn Riverside  
Newcastle upon Tyne  
NE15 8NY

**Responsible statistician:** `r config$responsible_statistician`

</main>