---
title: "`r config$publication_name`"
pagetitle: PCA England
always_allow_html: true
output:
  word_document:
    toc: yes
    toc_depth: '2'
  html_document:
    anchor_sections: no
    css: 
      - www/style.css
      - www/nhs-min.css
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
---

<script>

  
    $(document).ready(function() {
    $('#header').prepend('<img src=`r knitr::image_uri("www/Accredited Official Statistics Logo English.svg")` alt=\"NHSBSA logo\" style=\"position:absolute; top: 0; right: 0; padding: 10px; width: 25%; margin-top:140px;\">');
  });
</script>

<html lang="en">

```{r setup, include=FALSE}

# set code chunk options to disable echo by default
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE)


```


<div class = "quartoheader", id = "fixed-header">
  <header class="nhsuk-header" role="banner">
    <div class="nhsuk-width-container nhsuk-header__container">
      <div class="nhsuk-header__logo nhsuk-header__logo--only">
        <a class="nhsuk-header__link" href="https://www.nhsbsa.nhs.uk/" aria-label="NHSBSA home" target="_blank">
          <img class="nhsuk-logo" src="www/logo-nhsbsa.svg"" name="NHSBSA logo" alt="NHS Business Services Authority">
        </a>
      </div>
    </div>
</div>

<!-- JS to handle hiding the header on scroll -->
<script>
  document.addEventListener("scroll", function() {
    var header = document.getElementById("fixed-header");
    var toc = document.getElementById("TOC");
    var BackTopButton = document.getElementById("backtop-button");
    var scrollPosition = window.scrollY || document.documentElement.scrollTop;
  if(scrollPosition > 88) {
    header.style.top = "-200px";
    toc.style.setProperty('margin-top', '24px', 'important');
    BackTopButton.style.setProperty('visibility', 'visible', 'important');
  } else
  {
    header.style.top = "0";
    toc.style.setProperty('margin-top', '140px', 'important');
    BackTopButton.style.setProperty('visibility', 'hidden', 'important');
  }
}
);
</script>

<!-- JS to handle back to top button -->
<script>
       function BackTop() {
        document.body.scrollTop = 0;
        document.documentElement.scrollTop = 0;
    }
</script>

<button onclick="BackTop()" id="backtop-button" style="opacity: 0.7; pointer-events: initial; visibility: visible;"><span aria-hidden="true">▲</span>Back to top</button>


<main>
# `r config$publication_sub_name` {.toc-ignore}

`r paste("Published", config$publication_date)`

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-12} 
`r infoBox_border(header = "Feedback", text = paste0("We are collecting feedback about this publication. You can send us feedback using our ", htmltools::HTML(paste("<a href='",config$stats_survey_link,"'>Official Statistics feedback survey</a>")),"."), width = "100%")` 
::: 
::::

## Key findings

::::{.row style="display: flex; margin-bottom: 24px;"}
:::{.col-md-6} 
`r infoBox_no_border(header = "", text = htmltools::HTML("<b>In 2024/25, 1.26 billion prescription items were dispensed in the community in England at a cost of £11.2 billion.</b>"), width = "100%")` 
::: 
:::{.col-md-6} 
`r infoBox_no_border(header = "", text = htmltools::HTML("<b>The number of items dispensed in the community in England increased by 4% from 2023/24 while costs increased by 2%.</b>"), width = "100%")` 
:::
::::

In England in 2024/25 the most dispensed chemical substance was Atorvastatin, a drug used to lower cholesterol and prevent heart disease, with 73 million items. 

BNF Chapter 6 Endocrine System accounted for 19% of total costs in England in 2024/25.

BNF Chapter 2 Cardiovascular System accounted for 30% of all items dispensed in England in 2024/25.

Dapagliflozin, a medication mainly used to treat type 2 diabetes, was the chemical substance with the highest costs. In England in 2024/25, these costs were £333 million.

Forxiga 10mg tablets was the presentation with the largest increase in cost from 2023/24, with an increase of £97 million. Forxiga is a brand of the chemical substance dapagliflozin.

---

## 1. Things you should know

### 1.1. Scope {.toc-ignore}

Prescription Cost Analysis (PCA) provides information about the costs and volumes of prescription items that have been dispensed in England.

These statistics cover prescription items submitted to the NHSBSA for reimbursement that were prescribed in England, Scotland, Wales, Northern Ireland or the Channel Islands, and subsequently dispensed in the community in England. By ‘dispensed in the community’ we mean it was dispensed at a pharmacy, by an appliance contractor, by a dispensing doctor, or as a personally administered item. Items dispensed in hospitals or prisons, for example, are not described as ‘dispensed in the community’. Items prescribed in England but dispensed elsewhere in the United Kingdom are not included in these statistics.

The Pharmacy First Clinical Pathways advanced service was launched on 31 January 2024. The service allows patients to get treatment for seven common conditions directly from their local pharmacy, without the need for a GP appointment or prescription. This includes medicines to treat:

-	sinusitis
- sore throat
-	earache
-	infected insect bite
-	impetigo (a bacterial skin infection)
-	shingles
-	uncomplicated urinary tract infections in women

Data relating to medicines supplied as part of this advanced service have been included in this publication. 

More information in the Pharmacy First service can be found on the [NHS England website](https://www.england.nhs.uk/primary-care/pharmacy/pharmacy-services/pharmacy-first/).

This publication is an accredited official statistic release. Accredited official statistic status means that PCA meets the highest standards of trustworthiness, quality and public value, and complies with all aspects of the [Code of Practice for Statistics](https://code.statisticsauthority.gov.uk/).

The designation of this publication as a National Statistic was confirmed in May 2021 following an assessment by the Office for Statistics Regulation (OSR). More information about this confirmation, and the assessment of these statistics, can be found on the [OSR website](https://osr.statisticsauthority.gov.uk/correspondence/ed-humpherson-to-matthew-wilson-confirmation-of-national-statistics-designation-for-prescription-cost-analysis-england-statistics/).

### 1.2. Definitions {.toc-ignore}

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-12 .toc-ignore}
`r infoBox_border(header = "Item", text = "A single unit of medication listed separately on a prescription form. For example, in this publication, an item might be listed as paracetamol 500mg tablets x28, distinct from other medications that may be prescribed on the same form.", width = "100%")` 
::: 
::::

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-12} 
`r infoBox_border("NIC", text = "The Net Ingredient Cost (NIC), referred to as costs, is the basic price of a single unit of a medication multiplied by the quantity prescribed. It does not include other fees incurred by dispensing contractors, such as controlled drug fees or the single activity fee. The basic price is determined by the [Drug Tariff](https://www.nhsbsa.nhs.uk/pharmacies-gp-practices-and-appliance-contractors/drug-tariff) or by the manufacturer, wholesaler, or supplier of the product.", width = "100%")`
::: 
::::

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-12} 
`r infoBox_border("Classification", text = "This publication uses the British National Formulary (BNF), which lists medicines used in the UK and classifies them according to their primary therapeutic use. Medication may also be prescribed for medical conditions other than their primary use. For example, certain antidepressants may be recommended to individuals experiencing chronic pain. The NHSBSA does not capture the clinical indication of a prescription during processing.", width = "100%")`
::: 
::::

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-12} 
`r infoBox_border("Generic prescribing", text = "Generic prescribing is the prescribing of a drug by it’s non-proprietary name rather than a specific brand name. This allows for any suitable drug that meets the same standards as the brand name drug to be dispensed, which can result in cost savings. For example, paracetamol 500mg tablets.", width = "100%")`
::: 
::::

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-12} 
`r infoBox_border("Branded prescribing", text = "Branded prescribing is the prescribing of a drug by specific brand name. For example, Panadol Advance 500mg tablets.", width = "100%")`
::: 
::::

---

## 2. Results and commentary

### 2.1. Cost and number of prescription items dispensed in England 

#### `r paste0("Cost of prescription items dispensed in England 2015/16 to ", paste0(substr(max_data_fy,1,5), substr(max_data_fy, 8, 9)))` {.tabset}

<div class = "tabset">

##### Chart

###### Figure 1: The cost of prescription items dispensed in England has been increasing since 2018/19

Note: The y-axis does not start at zero. This is to emphasise relative changes rather than absolute values. Please consider the scale when interpreting the data.

```{r total_nic_chart}

figure_1
get_download_button(title = "Download chart data", data = figure_1_data, filename = "figure_1")

```

##### Table

###### Table 1: The cost of prescription items dispensed in England has been increasing since 2018/19

```{r total_nic_table}
knitr::kable(table_1, align = "lr")

get_download_button(title = "Download table data", data = figure_1_data, filename = "table_1")

```

</div>

Source: [Summary Tables - Prescription Cost Analysis - National](`r config$fy_nat_link`)

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "<b>Prescription items dispensed in the community in England in 2024/25 cost £11.2 billion.</b>", width = "100%")` 
::: 
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "<b>There was a 2% increase in costs from 2023/24 to 2024/25.</b>", width = "100%")` 
:::
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "<b>In England, the cost of prescription items increased 20% from 2015/16 to 2024/25, an increase of £1.86 billion.</b>", width = "100%")` 
:::
::::

#### `r paste0("Number of prescription items dispensed in England 2015/16 to ", paste0(substr(max_data_fy,1,5), substr(max_data_fy, 8, 9)))` {- .tabset}

<div class = "tabset">

##### Chart

###### Figure 2: The number of prescription items dispensed in England has been increasing since 2020/21

Note: The y-axis does not start at zero. This is to emphasise relative changes rather than absolute values. Please consider the scale when interpreting the data.

```{r total_items}

figure_2
get_download_button(title = "Download chart data", data = figure_2_data, filename = "figure_2")
```

##### Table

###### Table 2: The number of prescription items dispensed in England has been increasing since 2020/21

```{r total_items_table}
knitr::kable(table_2, align = "lr")

get_download_button(title = "Download table data", data = figure_2_data, filename = "table_2")

```

</div>

Source: [Summary Tables - Prescription Cost Analysis - National](`r config$fy_nat_link`)

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "<b>In 2024/25, 1.26 billion items were dispensed in the community in England.</b>", width = "100%")` 
::: 
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "<b>There was a 4% increase in items from 2023/24 to 2024/25.</b>", width = "100%")` 
:::
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "<b>In England, the number of prescription items dispensed increased by 16% from 2015/16 to 2024/25, an increase of 170 million items.</b>", width = "100%")` 
:::
::::

### 2.2. Cost and number of prescription items dispensed in England amongst BNF chapters

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-12} 
`r infoBox_border(header = "", text = "These statistics use the British National Formulary (BNF) to group medicines based on their primary therapeutic indication. The BNF has multiple levels, starting with the largest grouping called chapters, followed by sections, paragraphs, sub-paragraphs, chemical substances, products, and individual presentations.<br><br>The NHSBSA uses the BNF classification system implemented prior to the release of edition 70, which includes 6 additional chapters (18 to 23) created by NHS Prescription Services. These chapters are used to classify products that fall outside chapters 1 to 15, including dressings, appliances and medical devices.<br><br>However, presentations in chapters 20 to 23 do not have assigned BNF paragraphs, sub-paragraphs, chemical substances, or products.", width = "100%")` 
::: 
::::

#### `r paste0("Distribution of cost of items dispensed in England amongst BNF chapters ", paste0(substr(max_data_fy,1,5), substr(max_data_fy, 8, 9)))` {- .tabset}

<div class = "tabset">

##### Chart

###### Figure 3: In 2024/25, BNF Chapter 6 Endocrine System had the highest costs

```{r bnf_chapters_costs}

figure_3
get_download_button(title = "Download chart data", data = figure_3_data, filename = "figure_3")
```

##### Table

###### Table 3: In 2024/25, BNF Chapter 6 Endocrine System had the highest costs

```{r bnf_chapters_costs_table}
knitr::kable(table_3, align = "llr")

get_download_button(title = "Download table data", data = figure_3_data, filename = "table_3")

```

</div>

Source: [Summary Tables - Prescription Cost Analysis - Chapter](`r config$fy_nat_link`)

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "<b>BNF Chapter 6 Endocrine System accounted for 19% of total costs in 2024/25.</b>", width = "100%")` 
::: 
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "<b>BNF Chapter 4 Central Nervous System accounted for 15% of total costs in 2024/25.</b>", width = "100%")` 
:::
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "<b>BNF Chapter 2 Cardiovascular System accounted for 12% of total costs in 2024/25.</b>", width = "100%")`
:::
::::

In 2024/25, BNF Chapter 6 Endocrine System had the highest costs, totalling £2.07 billion. BNF Chapter 6 includes drugs used to treat diabetes and sex hormones. Chapter 4 Central Nervous System followed with costs of £1.69 billion. Chapter 2 Cardiovascular System had costs of £1.38 billion. The top 3 BNF Chapters by total costs remains consistent with 2023/24. However, Chapter 4 Central Nervous System has moved up to second place and BNF Chapter 2 Cardiovascular System down to third place.

#### Index of the BNF chapters with the largest percentage increase and decrease in costs from 2015/16 to 2024/25 {- .tabset}

<div class = "tabset">

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-12} 
`r infoBox_border(header = "", text = "These statistics only cover BNF Chapters 01 through to 15.", width = "100%")` 
::: 
::::

##### Chart

###### Figure 4: BNF Chapter 1 Gastro-Intestinal System has shown the largest percentage increase in costs from 2015/16 to 2024/25

```{r bnf_chapters_costs_index}
figure_4
get_download_button(title = "Download chart data", data = figure_4_data, filename = "figure_4")

```

##### Table

###### Table 4: BNF Chapter 1 Gastro-Intestinal System has shown the largest percentage increase in costs from 2015/16 to 2024/25

```{r bnf_chapters_costs_index_table}
knitr::kable(table_4, align = "lrr")

get_download_button(title = "Download table data", data = figure_4_data, filename = "table_4")

```

</div>

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-6} 
`r infoBox_no_border(header = "", text = "<b>BNF Chapter 1 Gastro-Intestinal System has shown the greatest percentage increase in costs from 2015/16 to 2024/25 of 55%.</b>", width = "100%")` 
::: 
:::{.col-md-6} 
`r infoBox_no_border(header = "", text = "<b>BNF Chapter 10 Musculoskeletal and Joint Diseases has shown the greatest percentage decrease in costs from 2015/16 to 2024/25 of 37%.</b>", width = "100%")` 
:::
::::

#### `r paste0("Distribution of items dispensed in England amongst BNF chapters ", paste0(substr(max_data_fy,1,5), substr(max_data_fy, 8, 9)))` {- .tabset}

<div class = "tabset">

##### Chart

###### Figure 5: In 2024/25, BNF Chapter 2 Cardiovascular System had the highest number of items

```{r bnf_chapters_items}

figure_5
get_download_button(title = "Download chart data", data = figure_5_data, filename = "figure_5")
```

##### Table

###### Table 5: In 2024/25, BNF Chapter 2 Cardiovascular System had the highest number of items

```{r bnf_chapters_items_table}
knitr::kable(table_5, align = "llr")

get_download_button(title = "Download table data", data = figure_5_data, filename = "table_5")

```

</div>

Source: [Summary Tables - Prescription Cost Analysis - Chapter](`r config$fy_nat_link`)

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "<b>BNF Chapter 2 Cardiovascular System accounted for 30% of all items dispensed in 2024/25.</b>", width = "100%")` 
::: 
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "<b>BNF Chapter 4 Central Nervous System accounted for 19% of all items dispensed in 2024/25.</b>", width = "100%")` 
:::
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "<b>BNF Chapter 6 Endocrine System accounted for 11% of all items dispensed in 2024/25.</b>", width = "100%")` 
:::
::::

In 2024/25, BNF Chapter 2 Cardiovascular System had the highest number of prescription items dispensed in England, with a total of 377 million items. Chapter 4 Central Nervous System followed with 238 million items. Chapter 6 Endocrine System had 141 million items. This mirrors the trend for the top 3 BNF Chapters by items dispensed as in 2023/24,

#### Index of the BNF chapters with the largest percentage increase and decrease in items from 2015/16 to 2024/25 {- .tabset}

<div class = "tabset">

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-12} 
`r infoBox_border(header = "", text = "These statistics only cover BNF Chapters 01 through to 15.", width = "100%")` 
::: 
::::

##### Chart

###### Figure 6: BNF Chapter 6 Endocrine System has shown the greatest percentage increase in items from 2015/16 to 2024/25

```{r bnf_chapters_items_index}
figure_6
get_download_button(title = "Download chart data", data = figure_6_data, filename = "figure_6")

```

##### Table

###### Table 6: BNF Chapter 6 Endocrine System has shown the greatest percentage increase in items from 2015/16 to 2024/25

```{r bnf_chapters_items_index_table}
knitr::kable(table_6, align = "lrr")

get_download_button(title = "Download table data", data = figure_6_data, filename = "table_6")

```

</div>

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-6} 
`r infoBox_no_border(header = "", text = "<b>BNF Chapter 6 Endocrine System has shown the greatest percentage increase in items from 2015/16 to 2024/25 of 36%.</b>", width = "100%")` 
::: 
:::{.col-md-6} 
`r infoBox_no_border(header = "", text = "<b>BNF Chapter 13 Skin has shown the greatest percentage decrease in items from 2015/16 to 2024/25 of 26%.</b>", width = "100%")` 
:::
::::

### 2.3. Generic prescribing and dispensing in England

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-12} 
`r infoBox_border(header = "", text = "Many drugs are now prescribed generically even when they are not available in generic form. For example, drugs that are listed in Category C of the Drug Tariff.", width = "100%")` 
::: 
::::

#### `r paste0("Proportion of items prescribed generically and proportion of costs excluding appliances England 2015/16 to ", paste0(substr(max_data_fy,1,5), substr(max_data_fy, 8, 9)))` {- .tabset}

<div class = "tabset">

##### Chart

###### Figure 7: Between 2015/16 and 2024/25 the percentage of items prescribed generically has remained consistent

Note: The y-axis does not start at zero. This is to emphasise relative changes rather than absolute values. Please consider the scale when interpreting the data.

```{r generic_presc}

figure_7
get_download_button(title = "Download chart data", data = figure_7_data, filename = "figure_7")
```

##### Table

###### Table 7: Between 2015/16 and 2024/25, the percentage of items prescribed generically has remained consistent

```{r generic_presc_table}
knitr::kable(table_7, align = "lrr")

get_download_button(title = "Download table data", data = figure_7_data, filename = "table_7")

```

</div>

Source: [Additional Tables - Prescription Cost Analysis - table A5](`r config$add_anl_link`)

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-6} 
`r infoBox_no_border(header = "", text = "<b>Generic items made up 86% of the total items prescribed in 2024/25.</b>", width = "100%")` 
::: 
:::{.col-md-6} 
`r infoBox_no_border(header = "", text = "<b>Generic prescribing accounted for 59% of costs in 2024/25.</b>", width = "100%")` 
:::
::::

The percentage of items prescribed generically has remained consistent between 2015/16 and 2024/25. There has been a 2 percentage point variation in the period between 2015/16 and 2024/25, meaning it has stayed stable over the period. Note that these figures exclude preparation class 4 products - appliances.

Between 2015/16 and 2024/25 the percentage of costs which was accounted for by generic prescribing fell by 8 percentage points.

#### `r paste0("Sankey chart to show the flow of prescribing and dispensing volumes England ", paste0(substr(max_data_fy,1,5), substr(max_data_fy, 8, 9)))` {- .tabset}

<div class = "tabset">

##### Chart

###### Figure 8: In 2024/25, the majority of items were both prescribed and dispensed generically

```{r presc_sankey, out.height=400}

figure_8
get_download_button(title = "Download chart data", data = figure_8_data, filename = "figure_8")
```

##### Table

###### Table 8: In 2024/25, the majority of items were both prescribed and dispensed generically

```{r presc_sankey_table}
knitr::kable(table_8, align = "llr")

get_download_button(title = "Download table data", data = figure_8_data, filename = "table_8")

```

</div>

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-6} 
`r infoBox_no_border(header = "", text = "<b>In 2024/25, 60 million of the 1.03 billion items prescribed generically were dispensed as a proprietary item, this is 6% of items prescribed generically.</b>", width = "100%")` 
::: 
:::{.col-md-6} 
`r infoBox_no_border(header = "", text = "<b>In 2024/25, 172 million items were prescribed proprietary. This is 14% of the 1.26 billion items dispensed.</b>", width = "100%")` 
:::
::::

### 2.4. The 10 most dispensed chemical substances by costs and items in England

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-12} 
`r infoBox_border(header = "", text = "Drugs containing the same active ingredient are referred to by the same name, regardless of presentation type or if the product is branded or generic. For example, Omeprazole 20mg and Omeprazole 40mg capsules both contain the same active ingredient. The name used to identify the active substance, Omeprazole in this example, is referred to in this publication as ‘chemical substance’. Products such as dressings and appliances in BNF Chapters 20 to 23 do not have a chemical substance assigned to them.", width = "100%")` 
::: 
::::

#### `r paste0("The 10 most dispensed chemical substances by cost in ", paste0(substr(max_data_fy,1,5), substr(max_data_fy, 8, 9)))` {- .tabset}

<div class = "tabset">

##### Chart

###### Figure 9: In England in 2024/25, dapagliflozin was the chemical substance with the highest cost

```{r top_10_costs}

figure_9
get_download_button(title = "Download chart data", data = figure_9_data, filename = "figure_9")
```

##### Table

###### Table 9: In England in 2024/25, dapagliflozin was the chemical substance with the highest cost

```{r top_10_costs_table}

knitr::kable(table_9, align = "llr")
get_download_button(title = "Download table data", data = figure_9_data, filename = "table_9")
```

</div>

Source: [Additional Tables - Prescription Cost Analysis - A2](`r config$add_anl_link`)

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "<b>Dapagliflozin had the highest cost in 2024/25 of £333 million. Dapagliflozin is a medication primarily used to treat type 2 diabetes. It is also used to treat adults with heart failure and chronic kidney disease.</b>", width = "100%")` 
::: 
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "<b>Three of the top 10 chemical substances by cost in 2024/25 are used to treat type 2 diabetes.</b>", width = "100%")` 
:::
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "<b>Two of the top 10 chemical substances by cost in 2024/25 are glucagon-like peptide-1 (GLP-1) receptor agonists used to help lose weight.</b>", width = "100%")` 
:::
::::

Dapagliflozin has risen from the third placed chemical substance by costs in England in 2023/24. Both of the GLP-1 receptor agonists, semglutide and tirzepatide, are new entries into the 10 most dispensed chemical substances by cost in 2024/25.

#### `r paste0("The 10 most dispensed chemical substances by total items in ", paste0(substr(max_data_fy,1,5), substr(max_data_fy, 8, 9)))` {- .tabset}

<div class = "tabset">

##### Chart

###### Figure 10: In England in 2024/25, Atorvastatin was the chemical substance with the highest number of items

```{r top_10_items}

figure_10
get_download_button(title = "Download chart data", data = figure_10_data, filename = "figure_10")
```

##### Table

###### Table 10: In England in 2024/25, Atorvastatin was the chemical substance with the highest number of items

```{r top_10_items_table}

knitr::kable(table_10, align = "llr")
get_download_button(title = "Download table data", data = figure_10_data, filename = "table_10")
```

</div>

Source: [Additional Tables - Prescription Cost Analysis - A3](`r config$add_anl_link`)

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "<b>Atorvastatin was the most dispensed chemical substance in 2024/25 with 73 million items. Atorvastatin is a type of statin commonly used to lower cholesterol and help prevent heart disease.</b>", width = "100%")` 
:::
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "<b>Atorvastatin has been the most dispensed chemical substance by items since 2016/17.</b>", width = "100%")` 
:::
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "<b>Four of the top 10 chemical substances by items in 2024/25 belong to BNF Chapter 2 Cardiovascular System. These substances are Atorvastatin, Amlodipine, Ramipril, and Bisoprolol fumarate.</b>", width = "100%")` 
:::
::::

### 2.5. Cost and number of prescription items dispensed by Integrated Care Board (ICB) in England

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-12} 
`r infoBox_border(header = "", text = paste("ICBs are statutory NHS organisations responsible for developing a plan for meeting the health needs of the population, managing the NHS budget, and arranging for the provision of health services in a geographical area. <br><br>It is important to note that the ICB in these statistics is the location of the dispensing contract as listed in NHSBSA administrative records, rather than the prescribing organisation of an item.<br><br>Prescribing and dispensing trends vary regionally, and there are many factors that influence these trends, such as the presence of specialist dispensing contractors within a given area.<br><br>To find out more about ICBs and how they were used in these statistics, please refer to section 3.2 of the [Background Information and Methodology note](",config$background_link,") that accompanies this release."), width = "100%")` 
::: 
::::

#### `r paste0("Cost per person (£) by ICB ",year_tidy)` {- .tabset}

<div class = "tabset">

##### Chart

###### Figure 11: The mean cost of items dispensed per person in England was £195 in 2024/25

```{r nic_per_pop}
figure_11
get_download_button(title = "Download chart data", data = figure_11_data, filename = "figure_11")
```

##### Table

###### Table 11: The mean cost of items dispensed per person in England was £195 in 2024/25

```{r nic_per_pop_table}
knitr::kable(table_11, align = "lrrr")
get_download_button(title = "Download table data", data = figure_11_data, filename = "table_11")
```

</div>

Source: [Summary Tables - Prescription Cost Analysis - ICB National](`r config$fy_icb_link`) | [Office for National Statistics - Health geographies population estimates](https://www.ons.gov.uk/file?uri=/peoplepopulationandcommunity/populationandmigration/populationestimates/datasets/clinicalcommissioninggroupmidyearpopulationestimates/mid2011tomid2022integratedcareboards2024geography/sapeicb202420112022.xlsx)

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-6} 
`r infoBox_no_border(header = "", text = "<b>NHS Cambridgeshire and Peterborough ICB continues to have the highest cost per person, in 2024/25 this was £417.</b>", width = "100%")` 
:::
:::{.col-md-6} 
`r infoBox_no_border(header = "", text = "<b>NHS South East London ICB continues to have the lowest cost per person, in 2024/25 this was £133.</b>", width = "100%")` 
:::
::::

As in 2023/24, NHS Cambridgeshire and Peterborough ICB's cost per person is more than double the national average. Dispensing contractors in this ICB offer specialist stoma care, 35% of their total costs of £387 million in 2024/25 were in BNF Chapter 23 Stoma Appliances.

#### `r paste0("Dispensed items per person by ICB ",year_tidy)` {- .tabset}

<div class = "tabset">

##### Chart

###### Figure 12: The mean number of items dispensed per person in England was 22 in 2024/25

```{r items_per_pop}

figure_12
get_download_button(title = "Download chart data", data = figure_12_data, filename = "figure_12")  
```

##### Table

###### Table 12: The mean number of items dispensed per person in England was 22 in 2024/25

```{r items_per_pop_table}
knitr::kable(table_12, align = "lrrr")
get_download_button(title = "Download table data", data = figure_12_data, filename = "table_12")
```

</div>

Source: [Summary Tables - Prescription Cost Analysis - ICB National](`r config$fy_icb_link`) | [Office for National Statistics - Mid-year population estimates](https://www.ons.gov.uk/file?uri=/peoplepopulationandcommunity/populationandmigration/populationestimates/datasets/clinicalcommissioninggroupmidyearpopulationestimates/mid2011tomid2022integratedcareboards2024geography/sapeicb202420112022.xlsx)

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-6} 
`r infoBox_no_border(header = "", text = "<b>NHS West Yorkshire ICB continues to have the highest dispensed items per person, in 2024/25 this was 35.</b>", width = "100%")` 
:::
:::{.col-md-6} 
`r infoBox_no_border(header = "", text = "<b>NHS Surrey Heartlands ICB continues to have the lowest dispensed items per person, in 2024/25 this was 14.</b>", width = "100%")` 
:::
::::

### 2.6. Top presentations by changes in unit and total costs

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-12} 
`r infoBox_border(header = "", text = "Presentation is the name given to the specific type, strength, and formulation of a drug, or the specific type of an appliance. For example, paracetamol 500mg tablets.<br><br>The charts in this section have been limited to presentations with a total cost in 2024/25 greater than £1 million.<br><br>A unit cost for a presentation is calculated by dividing the total cost of items dispensed in the period by the total quantity.<br><br>Price concessions can be granted by the Department of Health and Social Care (DHSC) for medicines that pharmacists are struggling to obtain at or below the price set out in the Drug Tariff. These concessions can have large impacts on the unit costs of drugs.", width = "100%")` 
::: 
::::

#### `r paste0("Top 10 presentations by percentage increase in unit cost ",paste0(substr(max_data_fy_minus_1,1,5), substr(max_data_fy_minus_1, 8, 9)), " to ", paste0(substr(max_data_fy,1,5), substr(max_data_fy, 8, 9)))` {- .tabset}

<div class = "tabset">

##### Chart

###### Figure 13: Hydrocortisone 0.5% cream had the highest increase in unit cost from 2023/24 to 2024/25

```{r unit_cost_increase}
figure_13
get_download_button(title = "Download chart data", data = figure_13_data, filename = "figure_13")
```

##### Table

###### Table 13: Hydrocortisone 0.5% cream had the highest increase in unit cost from 2023/24 to 2024/25

```{r unit_cost_increase_table}
knitr::kable(table_13, align = "llr")
get_download_button(title = "Download table data", data = figure_13_data, filename = "table_13")
```

</div>

Source: [Additional Tables - Prescription Cost Analysis - A11](`r config$add_anl_link`)

Hydrocortisone 0.5% cream had the highest increase in unit cost from 2023/24 to 2024/25, an increase of 2,241%. Hydrocortisone 0.5% cream is used to treat skin conditions such as eczema, psoriasis, and insect bites. Hydrocortisone 0.5% cream was removed from the Drug Tariff in March 2024.

#### `r paste0("Top 10 presentations by percentage decrease in unit cost ",paste0(substr(max_data_fy_minus_1,1,5), substr(max_data_fy_minus_1, 8, 9)), " to ", paste0(substr(max_data_fy,1,5), substr(max_data_fy, 8, 9)))`  {- .tabset}

<div class = "tabset">

##### Chart

###### Figure 14: Sitagliptin 100mg tablets had the highest decrease in unit cost from 2023/24 to 2024/25

```{r unit_cost_decrease}
figure_14
get_download_button(title = "Download chart data", data = figure_14_data, filename = "figure_14")
```

##### Table

###### Table 14: Sitagliptin 100mg tablets had the highest decrease in unit cost from 2023/24 to 2024/25

```{r unit_cost_decrease_table}
knitr::kable(table_14, align = "llr")
get_download_button(title = "Download table data", data = figure_14_data, filename = "table_14")
```

</div>

Source: [Additional Tables - Prescription Cost Analysis - A12](`r config$add_anl_link`)

Sitagliptin 100mg tablets had the highest decrease in unit cost from 2023/24 to 2024/25, a decrease of 86%. Sitagliptin came off patent and was made a generic in June 2023. Sitagliptin is used primarily to treat type 2 diabetes.

#### `r paste0("Top 10 presentations by increase in total cost ", paste0(substr(max_data_fy_minus_1,1,5), substr(max_data_fy_minus_1, 8, 9)), " to ", paste0(substr(max_data_fy,1,5), substr(max_data_fy, 8, 9)))` {- .tabset}

<div class = "tabset">

##### Chart

###### Figure 15: Forxiga 10mg tablets had the highest increase in total costs from 2023/24 to 2024/25

```{r abs_cost_increase}
figure_15
get_download_button(title = "Download chart data", data = figure_15_data, filename = "figure_15")
```

##### Table

###### Table 15: Forxiga 10mg tablets had the highest increase in total costs from 2023/24 to 2024/25

```{r abs_cost_increase_table}
knitr::kable(table_15, align = "llr")
get_download_button(title = "Download table data", data = figure_15_data, filename = "table_15")
```

</div>

Source: [Additional Tables - Prescription Cost Analysis - A13](`r config$add_anl_link`)

Between 2023/24 and 2024/25, the total costs of Forxiga 10mg tablets increased by £97 million, the highest increase of all presentations. This is the second year in a row where Forxiga 10mg tablets have shown the highest increase in costs following an increase of £90 million between 2022/23 and 2023/24. Forxiga 10mg tablets is the proprietary version of Dapagliflozin 10mg tablets, primarily used to control blood sugar levels in people with type 2 diabetes. 

#### `r paste0("Top 10 presentations by decrease in total cost ", paste0(substr(max_data_fy_minus_1,1,5), substr(max_data_fy_minus_1, 8, 9)), " to ", paste0(substr(max_data_fy,1,5), substr(max_data_fy, 8, 9)))` {- .tabset}

<div class = "tabset">

##### Chart

###### Figure 16: Apixaban 5mg tablets had the highest decrease in total costs from 2023/24 to 2024/25

```{r abs_cost_decrease}
figure_16
get_download_button(title = "Download chart data", data = figure_16_data, filename = "figure_16")
```

##### Table

###### Table 16: Apixaban 5mg tablets had the highest decrease in total costs from 2023/24 to 2024/25

```{r abs_cost_decrease_table}
knitr::kable(table_16, align = "llr")
get_download_button(title = "Download table data", data = figure_16_data, filename = "table_16")
```

</div>

Source: [Additional Tables - Prescription Cost Analysis - A14](`r config$add_anl_link`)

Between 2023/24 and 2024/25, the total costs of Apixaban 5mg tablets decreased by £131 million, the highest decrease of all presentations. This is the second year in a row where Apixaban 5mg tablets have shown the highest decrease in costs following a decrease of £120 million between 2022/23 and 2023/24. A generic version of Apixiban was made available in May 2022. Apixiban is an anticoagulant primarily used to lower the risk of blood clots developing in patients who have a condition called atrial fibrillation.

---

## 3. UK comparisons

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-12} 
`r infoBox_border(header = "", text = paste("Prescription Cost Analysis statistics are produced separately by each UK nation, and there are differences in data collection, healthcare structures, and methodology. Therefore, it's important to be cautious when comparing PCA data of different UK nations, and to consider the broader context of the UK's health and social care landscape.<br><br>We have chosen a few high-level measures that allow for broad comparisons across the devolved administrations. For more detailed information, please refer to each country's official PCA publication<br><br>Our comparisons are based on the most recent PCA data available at the time of analysis and use the Office for National Statistics (ONS) 2023 mid-year population estimates to calculate per-person figures for the 2023/24 financial year.<br><br>For Wales, the figures reflect prescriptions issued by Welsh GPs and dispensed either in Wales or elsewhere in the UK.<br><br>For Northern Ireland the PCA data used is for the calendar year 2024."), width = "100%")` 
::: 
::::

#### `r paste0("Cost per person for the UK nations ", max_data_fy_minus_1)` {- .tabset}

<div class = "tabset">

##### Chart

###### Figure 17: England had the lowest prescription item cost per person in 2023/24

```{r uk_comparisons_costs}
figure_17
get_download_button(title = "Download chart data", data = figure_17_data, filename = "figure_17")

```

##### Table

###### Table 17: England had the lowest prescription item cost per person in 2023/24

```{r uk_comparisons_costs_table}
knitr::kable(table_17, align = "lr")
get_download_button(title = "Download table data", data = figure_17_data, filename = "table_17")
```

</div>

Source: [Additional Tables - Prescription Cost Analysis - A1](`r config$add_anl_link`) | [Public Health Scotland - Dispenser payments and prescription cost analysis](`r config$scotland_pca`) | [Welsh Government - Prescriptions in Wales](`r config$wa_pca_link`) | [HSC Business Services Organisation - General pharmaceutical services and prescribing statistics](`r config$ni_pca_link`) | [Office for National Statistics - Mid-year population estimates](https://www.ons.gov.uk/peoplepopulationandcommunity/populationandmigration/populationestimates/datasets/populationestimatestimeseriesdataset)

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-6} 
`r infoBox_no_border(header = "", text = "<b>England had the lowest prescription item cost per person in 2023/24 at £189.</b>", width = "100%")` 
::: 
:::{.col-md-6} 
`r infoBox_no_border(header = "", text = "<b>Northern Ireland had the highest prescription item cost per person in 2023/24 at £261</b>", width = "100%")` 
::: 
::::

#### `r paste0("Items per person for the UK nations ", max_data_fy_minus_1)` {- .tabset}

<div class = "tabset">

##### Chart

###### Figure 18: Scotland had the fewest items dispensed per person in 2023/24

```{r uk_comparisons_items}

figure_18
get_download_button(title = "Download chart data", data = figure_18_data, filename = "figure_18")
```

##### Table

###### Table 18: Scotland had the fewest items dispensed per person in 2023/24

```{r uk_comparisons_items_table}
knitr::kable(table_18, align = "lr")
get_download_button(title = "Download table data", data = figure_18_data, filename = "table_18")
```

</div>

Source: [Additional Tables - Prescription Cost Analysis - A1](`r config$add_anl_link`) | [Public Health Scotland - Dispenser payments and prescription cost analysis](`r config$scotland_pca`) | [Welsh Government - Prescriptions in Wales](`r config$wa_pca_link`) | [HSC Business Services Organisation - General pharmaceutical services and prescribing statistics](`r config$ni_pca_link`) | [Office for National Statistics - Mid-year population estimates](https://www.ons.gov.uk/peoplepopulationandcommunity/populationandmigration/populationestimates/datasets/populationestimatestimeseriesdataset)

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-6} 
`r infoBox_no_border(header = "", text = "<b>Scotland had the lowest number of prescription items dispensed per person in 2023/24 at 21 items.</b>", width = "100%")` 
::: 
:::{.col-md-6} 
`r infoBox_no_border(header = "", text = "<b>Wales had the highest number of prescription items dispensed per person in 2023/24 at 27 items.</b>", width = "100%")` 
::: 
::::


---

## 4. About these statistics

### 4.1. Monthly PCA data {.toc-ignore}

The NHSBSA publishes a [monthly administrative management information dataset for PCA](`r config$nhsbsa_pca_link`) using the same methodology as these statistics. However, while our annual PCA is an Official Statistic, the monthly dataset is not.

There is a slight difference between the annual National Statistics and the monthly administrative data for PCA. The annual publication uses the geographical structure as of the 31 March for that year, while the monthly data uses the structure at the end of the month. If a dispensing contractor changes parent organisation during the year, this may affect the ICB level total in the aggregated monthly data compared to the annual National Statistics, but it will not affect the national totals. However, such cases are rare.


### 4.2. Uses of these statistics {.toc-ignore}

These statistics serve several purposes, including:

* providing a national overview of community prescription costs and volumes in England for both the financial year and the calendar year

* analysing cost and volume trends across different levels of the BNF hierarchy

* allowing public scrutiny of national prescribing habits

* monitoring uptake of new to market proprietary presentations

* monitoring the proportion of prescriptions dispensed generically

### 4.3. Changes to this publication {.toc-ignore}

For this release, we have included supporting tables aggregated by NHS region.

The supporting tables now include population figures at all levels. These have been used to provide a measure of 'Items Per 1,000 Population' at all levels of data.

Information on items dispensed via the [Pharmacy First Clinical Pathway](https://www.england.nhs.uk/primary-care/pharmacy/pharmacy-services/pharmacy-first/) is included in the BNF presentation level data and SNOMED level data tables of the supplementary tables.

For full details of the changes made please refer to the [Background Information and Methodology note](`r config$background_link`) released alongside these statistics.

### 4.4. Timeliness of this publication {.toc-ignore}

This publication is released in June of each year, when March data becomes available. This allows us to provide a full financial year overview. Calendar year data is published in February of each year when December data becomes available on the [NHSBSA Open Data Portal](https://opendata.nhsbsa.net/) (ODP).

### 4.5. Price concessions {.toc-ignore}

When community pharmacies cannot source a drug at or below the reimbursement price as set out in the Drug Tariff, the DHSC can introduce a [price concession](https://cpe.org.uk/funding-and-reimbursement/reimbursement/price-concessions/) at the request of Community Pharmacy England. A price concession can be requested for any drugs listed in Part VIIIA, Part VIIIB, and Part VIIID of the Drug Tariff. For any drugs granted price concessions, contractors are automatically reimbursed at the new prices for that month.

---

## 5. Rounding

The high-level figures in this statistical summary have been rounded as per the table below:

```{r rounding_table}
rounding_table <- data.table(
  "From" = c("0", "1,001", "10,001", "100,001", "1,000,001", "10,000,001", "100,000,001", "10,000,000,001"),
  "To" = c("1,000", "10,000", "100,000", "1,000,000", "10,000,000", "100,000,000", "10,000,000,000", "100,000,000,000"),
  "Round to nearest" = c("1", "100", "1,000", "10,000", "100,000", "1,000,000", "10,000,000", "100,000,000")
)

knitr::kable(rounding_table, align = "rrr")
```

All changes and totals are calculated prior to rounding. Percentage changes are calculated prior to rounding and then are rounded to the nearest whole number. Since all figures within this statistical summary have been rounded, they may not match totals elsewhere when aggregated.

The summary tables released with this publication allow users to investigate this data at lower levels of granularity. Figures in the supplementary tables have not been rounded.

---

## 6. Accessibility

If you need information on this website in a different format, such as accessible PDF, large print, easy read, audio recording or braille, you can contact us by:

**Email**: accessibility@nhsbsa.nhs.uk

**Telephone**: 0191 203 5318

[Find out about call charges](https://www.nhsbsa.nhs.uk/contact-us/call-charges-and-phone-numbers)

We will consider your request and get back to you in 5 working days.

These contact details are only for accessibility queries. This email address is not for technical queries or IT problems. If you have a query that is not about accessibility, go to the ‘Feedback and contact us’ section of this page.

Read our [Accessibility statement for Official Statistics Narratives](https://www.nhsbsa.nhs.uk/accessibility-statement-official-statistics-narratives).

---

## 7. Feedback and contact us

Feedback is important to us. We welcome any questions and comments relating to these statistics. 

You can complete a [short survey about this publication](`r config$stats_survey_link`) to help us improve the Official Statistics that we produce. All responses will remain anonymous, and individuals will not be identifiable in any report that we produce.

You can view our [privacy policy](https://www.nhsbsa.nhs.uk/our-policies/privacy) on our website to see how your data is used and stored.

You can contact us by:

**Email:** statistics@nhsbsa.nhs.uk

**You can also write to us at:**

NHSBSA - Statistics  
NHS Business Services Authority  
Stella House  
Goldcrest Way  
Newburn Riverside  
Newcastle upon Tyne  
NE15 8NY

**Responsible statistician:** `r config$responsible_statistician`

</main>