---
title: "`r config$publication_name`"
pagetitle: PCA England
always_allow_html: true
output:
  word_document:
    toc: yes
    toc_depth: '2'
  html_document:
    anchor_sections: no
    css: 
      - www/style.css
      - www/nhs-min.css
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
---

<script>

  
    $(document).ready(function() {
    $('#header').prepend('<img src=`r knitr::image_uri("www/Accredited Official Statistics Logo English.svg")` alt=\"NHSBSA logo\" style=\"position:absolute; top: 0; right: 0; padding: 10px; width: 25%; margin-top:140px;\">');
  });
</script>

<html lang="en">

```{r setup, include=FALSE}

# set code chunk options to disable echo by default
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE)


```


<div class = "quartoheader", id = "fixed-header">
  <header class="nhsuk-header" role="banner">
    <div class="nhsuk-width-container nhsuk-header__container">
      <div class="nhsuk-header__logo nhsuk-header__logo--only">
        <a class="nhsuk-header__link" href="https://www.nhsbsa.nhs.uk/" aria-label="NHSBSA home" target="_blank">
          <img class="nhsuk-logo" src="www/logo-nhsbsa.svg"" name="NHSBSA logo" alt="NHS Business Services Authority">
        </a>
      </div>
    </div>
</div>

<!-- JS to handle hiding the header on scroll -->
<script>
  document.addEventListener("scroll", function() {
    var header = document.getElementById("fixed-header");
    var toc = document.getElementById("TOC");
    var BackTopButton = document.getElementById("backtop-button");
    var scrollPosition = window.scrollY || document.documentElement.scrollTop;
  if(scrollPosition > 88) {
    header.style.top = "-200px";
    toc.style.setProperty('margin-top', '24px', 'important');
    BackTopButton.style.setProperty('visibility', 'visible', 'important');
  } else
  {
    header.style.top = "0";
    toc.style.setProperty('margin-top', '140px', 'important');
    BackTopButton.style.setProperty('visibility', 'hidden', 'important');
  }
}
);
</script>

<!-- JS to handle back to top button -->
<script>
       function BackTop() {
        document.body.scrollTop = 0;
        document.documentElement.scrollTop = 0;
    }
</script>

<button onclick="BackTop()" id="backtop-button" style="opacity: 0.7; pointer-events: initial; visibility: visible;"><span aria-hidden="true">▲</span>Back to top</button>


<main>
# `r config$publication_sub_name` {.toc-ignore}

`r paste("Published", config$publication_date)`

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-12} 
`r infoBox_border(header = "Feedback", text = paste0("We are collecting feedback about this publication. You can send us feedback using our ", htmltools::HTML(paste("<a href='",config$stats_survey_link,"'>Official Statistics feedback survey</a>")),"."), width = "100%")` 
::: 
::::

## Key findings

In England in 2025/26:

::::{.row style="display: flex; margin-bottom: 24px;"}
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = htmltools::HTML("<b>1.30 billion prescription items were dispensed in the community at a cost of £11.6 billion.</b>"), width = "100%")` 
::: 
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = htmltools::HTML("<b>The number of items dispensed in the community increased by 3% from 2024/25 while costs increased by 4%.</b>"), width = "100%")` 
:::
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = htmltools::HTML("<b>The most dispensed chemical substance was Atorvastatin, a drug used to lower cholesterol and prevent heart disease, with 78 million items. </b>"), width = "100%")` 
:::
::::

---

## 1. Things you should know

### 1.1. Scope {.toc-ignore}

Prescription Cost Analysis (PCA) provides information about the costs and volumes of prescription items that have been dispensed in England.

These statistics cover prescription items submitted to the NHSBSA for reimbursement that were prescribed in England, Scotland, Wales, Northern Ireland or the Channel Islands, and subsequently dispensed in the community in England. By ‘dispensed in the community’ we mean it was dispensed at a pharmacy, by an appliance contractor, by a dispensing doctor, or as a personally administered item. Items dispensed in hospitals or prisons, for example, are not described as ‘dispensed in the community’. Items prescribed in England but dispensed elsewhere in the United Kingdom are not included in these statistics.

Pharmacy contractors are able to issue certain medicines through advanced services that they provide to NHS patients, data relating to these services are included in these statistics.

#### 1.1.1. Pharmacy First Clinical Pathways

The Pharmacy First Clinical Pathways advanced service was launched on 31 January 2024. The service allows patients to get treatment for seven common conditions directly from their local pharmacy, without the need for a GP appointment or prescription. This includes medicines to treat:

-	sinusitis
- sore throat
-	earache
-	infected insect bite
-	impetigo (a bacterial skin infection)
-	shingles
-	uncomplicated urinary tract infections in women

More information in the Pharmacy First service can be found on the [NHS England website](https://www.england.nhs.uk/primary-care/pharmacy/pharmacy-services/pharmacy-first/).

#### 1.1.2. NHS Pharmacy Contraception Service (PCS)

The PCS:  

- offers people greater choice on accessing contraception services  
- helps address health inequalities by providing wider healthcare access for high-risk communities and vulnerable patients in their communities and signposting service users to local sexual health services
- creates extra capacity in primary care and sexual health clinics (or equivalent) to support meeting the demand for more complex assessments 

More information in the PCS can be found on the [NHS England website](https://www.england.nhs.uk/primary-care/pharmacy/pharmacy-services/nhs-pharmacy-contraception-service/).

#### 1.1.3. Seasonal Influenza Vaccination Service

The Seasonal Influenza Vaccination Advanced Service allows eligible patients in England to receive a free NHS flu jab at their local pharmacy. Commissioned by NHS England, it provides a convenient alternative to GP appointments for at-risk groups.

More information in the Seasonal Influenza Vaccination Service can be found on the [NHS England website](https://www.england.nhs.uk/wp-content/uploads/2025/07/prn01998-v-community-pharmacy-advanced-service-specification-seasonal-influenza-vaccination-1-sept-2025-31-march-2026.pdf).

This publication is an accredited official statistic release. Accredited official statistic status means that PCA meets the highest standards of trustworthiness, quality and public value, and complies with all aspects of the [Code of Practice for Statistics](https://code.statisticsauthority.gov.uk/).

The designation of this publication as a National Statistic was confirmed in May 2021 following an assessment by the Office for Statistics Regulation (OSR). More information about this confirmation, and the assessment of these statistics, can be found on the [OSR website](https://osr.statisticsauthority.gov.uk/correspondence/ed-humpherson-to-matthew-wilson-confirmation-of-national-statistics-designation-for-prescription-cost-analysis-england-statistics/).

### 1.2. Definitions {.toc-ignore}

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-12 .toc-ignore}
`r infoBox_border(header = "Item", text = "A single unit of medication listed separately on a prescription form. For example, in this publication, an item might be listed as paracetamol 500mg tablets x28, distinct from other medications that may be prescribed on the same form.", width = "100%")` 
::: 
::::

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-12} 
`r infoBox_border("NIC", text = "The Net Ingredient Cost (NIC), referred to as costs, is the basic price of a single unit of a medication multiplied by the quantity prescribed. It does not include other fees incurred by dispensing contractors, such as controlled drug fees or the single activity fee. The basic price is determined by the [Drug Tariff](https://www.nhsbsa.nhs.uk/pharmacies-gp-practices-and-appliance-contractors/drug-tariff) or by the manufacturer, wholesaler, or supplier of the product.", width = "100%")`
::: 
::::

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-12} 
`r infoBox_border("Classification", text = "This publication uses the British National Formulary (BNF), which lists medicines used in the UK and classifies them according to their primary therapeutic use. Medication may also be prescribed for medical conditions other than their primary use. For example, certain antidepressants may be recommended to individuals experiencing chronic pain. The NHSBSA does not capture the clinical indication of a prescription during processing.", width = "100%")`
::: 
::::

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-12} 
`r infoBox_border("Generic prescribing", text = "Generic prescribing is the prescribing of a drug by it’s non-proprietary name rather than a specific brand name. This allows for any suitable drug that meets the same standards as the brand name drug to be dispensed, which can result in cost savings. For example, paracetamol 500mg tablets.", width = "100%")`
::: 
::::

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-12} 
`r infoBox_border("Branded prescribing", text = "Branded prescribing is the prescribing of a drug by specific brand name. For example, Panadol Advance 500mg tablets.", width = "100%")`
::: 
::::

---

## 2. Results and commentary

### 2.1. Cost and number of prescription items dispensed in England 

#### `r paste0("Cost of prescription items dispensed in England 2016/17 to ", paste0(substr(max_data_fy,1,5), substr(max_data_fy, 8, 9)))` {.tabset}

<div class = "tabset">

##### Chart

###### Figure 1: The cost of prescription items dispensed in England has been increasing since 2018/19

Note: The y-axis does not start at zero. This is to emphasise relative changes rather than absolute values. Please consider the scale when interpreting the data.

```{r total_nic_chart}

figure_1
get_download_button(title = "Download chart data", data = figure_1_data, filename = "figure_1")

```

##### Table

###### Table 1: The cost of prescription items dispensed in England has been increasing since 2018/19

```{r total_nic_table}
knitr::kable(table_1, align = "lr")

get_download_button(title = "Download table data", data = figure_1_data, filename = "table_1")

```

</div>

Source: [Summary Tables - Prescription Cost Analysis - National](`r config$fy_nat_link`)

In England in 2025/26:

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "<b>Prescription items dispensed in the community cost £11.6 billion.</b>", width = "100%")` 
::: 
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "<b>There was a 4% increase in costs from 2024/25.</b>", width = "100%")` 
:::
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "<b>The cost of prescription items increased 27% from 2016/17, an increase of £2.4 billion.</b>", width = "100%")` 
:::
::::

#### `r paste0("Number of prescription items dispensed in England 2016/17 to ", paste0(substr(max_data_fy,1,5), substr(max_data_fy, 8, 9)))` {- .tabset}

<div class = "tabset">

##### Chart

###### Figure 2: The number of prescription items dispensed in England has been increasing since 2020/21

Note: The y-axis does not start at zero. This is to emphasise relative changes rather than absolute values. Please consider the scale when interpreting the data.

```{r total_items}

figure_2
get_download_button(title = "Download chart data", data = figure_2_data, filename = "figure_2")
```

##### Table

###### Table 2: The number of prescription items dispensed in England has been increasing since 2020/21

```{r total_items_table}
knitr::kable(table_2, align = "lr")

get_download_button(title = "Download table data", data = figure_2_data, filename = "table_2")

```

</div>

Source: [Summary Tables - Prescription Cost Analysis - National](`r config$fy_nat_link`)

In England in 2025/26:

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "<b>There were 1.30 billion items dispensed in the community.</b>", width = "100%")` 
::: 
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "<b>There was a 3% increase in items  dispensed in the community from 2024/25.</b>", width = "100%")` 
:::
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "<b>The number of prescription items dispensed increased by 17% from 2016/17, an increase of 190 million items.</b>", width = "100%")` 
:::
::::

### 2.2. Cost and number of prescription items dispensed in England amongst BNF chapters

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-12} 
`r infoBox_border(header = "", text = "These statistics use the British National Formulary (BNF) to group medicines based on their primary therapeutic indication. The BNF has multiple levels, starting with the largest grouping called chapters, followed by sections, paragraphs, sub-paragraphs, chemical substances, products, and individual presentations.<br><br>The NHSBSA uses the BNF classification system implemented prior to the release of edition 70, which includes 6 additional chapters (18 to 23) created by NHS Prescription Services. These chapters are used to classify products that fall outside chapters 1 to 15, including dressings, appliances and medical devices.<br><br>However, presentations in chapters 20 to 23 do not have assigned BNF paragraphs, sub-paragraphs, chemical substances, or products.", width = "100%")` 
::: 
::::

#### `r paste0("Distribution of cost of items dispensed in England amongst BNF chapters ", paste0(substr(max_data_fy,1,5), substr(max_data_fy, 8, 9)))` {- .tabset}

<div class = "tabset">

##### Chart

###### Figure 3: In 2025/26, BNF Chapter 6 Endocrine System had the highest costs

```{r bnf_chapters_costs}

figure_3
get_download_button(title = "Download chart data", data = figure_3_data, filename = "figure_3")
```

##### Table

###### Table 3: In 2025/26, BNF Chapter 6 Endocrine System had the highest costs

```{r bnf_chapters_costs_table}
knitr::kable(table_3, align = "llr")

get_download_button(title = "Download table data", data = figure_3_data, filename = "table_3")

```

</div>

Source: [Summary Tables - Prescription Cost Analysis - Chapter](`r config$fy_nat_link`)

In England in 2025/26:

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "<b>BNF Chapter 6 Endocrine System accounted for 21% of total costs.</b>", width = "100%")` 
::: 
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "<b>BNF Chapter 4 Central Nervous System accounted for 14% of total costs.</b>", width = "100%")` 
:::
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "<b>BNF Chapter 3 Respiratory System accounted for 11% of total costs.</b>", width = "100%")`
:::
::::

The share of costs attributed to BNF Chapter 6 Endocrine System increased by 2 percentage points, rising from 19% in 2024/25.

#### Index of the BNF chapters with the largest percentage increase and decrease in costs from 2016/17 to 2025/26 {- .tabset}

<div class = "tabset">

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-12} 
`r infoBox_border(header = "", text = "These statistics only cover BNF Chapters 01 through to 15.", width = "100%")` 
::: 
::::

##### Chart

###### Figure 4: BNF Chapter 14 Immunological Products and Vaccines has shown the largest percentage increase in costs from 2016/17 to 2025/26

```{r bnf_chapters_costs_index}
figure_4
get_download_button(title = "Download chart data", data = figure_4_data, filename = "figure_4")

```

##### Table

###### Table 4: BNF Chapter 14 Immunological Products and Vaccines has shown the largest percentage increase in costs from 2016/17 to 2025/26

```{r bnf_chapters_costs_index_table}
knitr::kable(table_4, align = "lrr")

get_download_button(title = "Download table data", data = figure_4_data, filename = "table_4")

```

</div>

In England in 2025/26:

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-6} 
`r infoBox_no_border(header = "", text = "<b>BNF Chapter 14 Immunological Products and Vaccines has shown the greatest percentage increase in costs from 2016/17 of 119%.</b>", width = "100%")` 
::: 
:::{.col-md-6} 
`r infoBox_no_border(header = "", text = "<b>BNF Chapter 10 Musculoskeletal and Joint Diseases has shown the greatest percentage decrease in costs from 2016/17 of 25%.</b>", width = "100%")` 
:::
::::

This publication now includes data on items issued under the Seasonal Influenza Vaccination Service, these items all fall under BNF Chapter 14. This combined with changes in the types of products issued and price increase has caused the jump in costs in BNF Chapter 14 seen between 2024/25 and 2025/26.

#### `r paste0("Distribution of items dispensed in England amongst BNF chapters ", paste0(substr(max_data_fy,1,5), substr(max_data_fy, 8, 9)))` {- .tabset}

<div class = "tabset">

##### Chart

###### Figure 5: In 2025/26, BNF Chapter 2 Cardiovascular System had the highest number of items

```{r bnf_chapters_items}

figure_5
get_download_button(title = "Download chart data", data = figure_5_data, filename = "figure_5")
```

##### Table

###### Table 5: In 2025/26, BNF Chapter 2 Cardiovascular System had the highest number of items

```{r bnf_chapters_items_table}
knitr::kable(table_5, align = "llr")

get_download_button(title = "Download table data", data = figure_5_data, filename = "table_5")

```

</div>

Source: [Summary Tables - Prescription Cost Analysis - Chapter](`r config$fy_nat_link`)

In England in 2025/26:

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "<b>BNF Chapter 2 Cardiovascular System accounted for 30% of all items dispensed.</b>", width = "100%")` 
::: 
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "<b>BNF Chapter 4 Central Nervous System accounted for 19% of all items dispensed.</b>", width = "100%")` 
:::
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "<b>BNF Chapter 6 Endocrine System accounted for 11% of all items dispensed.</b>", width = "100%")` 
:::
::::

The proportion of items dispensed in England from the top three BNF chapters in 2025/26 remained unchanged from 2024/25.

#### Index of the BNF chapters with the largest percentage increase and decrease in items from 2015/16 to 2024/25 {- .tabset}

<div class = "tabset">

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-12} 
`r infoBox_border(header = "", text = "These statistics only cover BNF Chapters 01 through to 15.", width = "100%")` 
::: 
::::

##### Chart

###### Figure 6: BNF Chapter 6 Endocrine System has shown the greatest percentage increase in items from 2016/17 to 2025/26

```{r bnf_chapters_items_index}
figure_6
get_download_button(title = "Download chart data", data = figure_6_data, filename = "figure_6")

```

##### Table

###### Table 6: BNF Chapter 6 Endocrine System has shown the greatest percentage increase in items from 2016/17 to 2025/26

```{r bnf_chapters_items_index_table}
knitr::kable(table_6, align = "lrr")

get_download_button(title = "Download table data", data = figure_6_data, filename = "table_6")

```

</div>

In England in 2025/26:

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-6} 
`r infoBox_no_border(header = "", text = "<b>BNF Chapter 6 Endocrine System has shown the greatest percentage increase in items from 2016/17 to 2025/26 of 37%.</b>", width = "100%")` 
::: 
:::{.col-md-6} 
`r infoBox_no_border(header = "", text = "<b>BNF Chapter 13 Skin has shown the greatest percentage decrease in items from 2016/17 to 2025/26 of 25%.</b>", width = "100%")` 
:::
::::

### 2.3. Generic prescribing and dispensing in England

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-12} 
`r infoBox_border(header = "", text = "Many drugs are now prescribed generically even when they are not available in generic form. For example, drugs that are listed in Category C of the Drug Tariff.<br><br>Appliances are excluded from analyses of generic prescribing because they do not have generic equivalents.", width = "100%")` 
::: 
::::

#### `r paste0("Proportion of items prescribed generically and proportion of costs excluding appliances England 2016/17 to ", paste0(substr(max_data_fy,1,5), substr(max_data_fy, 8, 9)))` {- .tabset}

<div class = "tabset">

##### Chart

###### Figure 7: Between 2016/17 and 2025/26 the percentage of items prescribed generically has remained consistent

Note: The y-axis does not start at zero. This is to emphasise relative changes rather than absolute values. Please consider the scale when interpreting the data.

```{r generic_presc}

figure_7
get_download_button(title = "Download chart data", data = figure_7_data, filename = "figure_7")
```

##### Table

###### Table 7: Between 2016/17 and 2025/26, the percentage of items prescribed generically has remained consistent

```{r generic_presc_table}
knitr::kable(table_7, align = "lrr")

get_download_button(title = "Download table data", data = figure_7_data, filename = "table_7")

```

</div>

Source: [Additional Tables - Prescription Cost Analysis - table A5](`r config$add_anl_link`)

In England in 2025/26:

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-6} 
`r infoBox_no_border(header = "", text = "<b>Generic items made up 86% of the total items prescribed.</b>", width = "100%")` 
::: 
:::{.col-md-6} 
`r infoBox_no_border(header = "", text = "<b>Generic prescribing accounted for 56% of costs.</b>", width = "100%")` 
:::
::::

The percentage of items prescribed generically has remained consistent between 2016/17 and 2025/26. There has been a 2 percentage point variation in the period, meaning it has stayed stable over the period. Note that these figures exclude preparation class 4 products - appliances.

Between 2016/17 and 2025/26 the percentage of costs which was accounted for by generic prescribing fell by 9 percentage points.

#### `r paste0("Sankey chart to show the flow of prescribing and dispensing volumes England ", paste0(substr(max_data_fy,1,5), substr(max_data_fy, 8, 9)))` {- .tabset}

<div class = "tabset">

##### Chart

###### Figure 8: In 2025/26, the majority of items were both prescribed and dispensed generically

```{r presc_sankey, out.height=400}

figure_8
get_download_button(title = "Download chart data", data = figure_8_data, filename = "figure_8")
```

##### Table

###### Table 8: In 2025/26, the majority of items were both prescribed and dispensed generically

```{r presc_sankey_table}
knitr::kable(table_8, align = "llr")

get_download_button(title = "Download table data", data = figure_8_data, filename = "table_8")

```

</div>

In England in 2025/26:

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-6} 
`r infoBox_no_border(header = "", text = "<b>48 million of the 1.06 billion items prescribed generically were dispensed as a proprietary item, this is 4% of items prescribed generically.</b>", width = "100%")` 
::: 
:::{.col-md-6} 
`r infoBox_no_border(header = "", text = "<b>171 million items were prescribed proprietary.</b>", width = "100%")` 
:::
::::

### 2.4. The 10 most dispensed chemical substances by costs and items in England

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-12} 
`r infoBox_border(header = "", text = "Drugs containing the same active ingredient are referred to by the same name, regardless of presentation type or if the product is branded or generic. For example, Omeprazole 20mg and Omeprazole 40mg capsules both contain the same active ingredient. The name used to identify the active substance, Omeprazole in this example, is referred to in this publication as ‘chemical substance’. Products such as dressings and appliances in BNF Chapters 20 to 23 do not have a chemical substance assigned to them.", width = "100%")` 
::: 
::::

#### `r paste0("The 10 most dispensed chemical substances by cost in ", paste0(substr(max_data_fy,1,5), substr(max_data_fy, 8, 9)))` {- .tabset}

<div class = "tabset">

##### Chart

###### Figure 9: In England in 2025/26, tirzepatide was the chemical substance with the highest cost

```{r top_10_costs}

figure_9
get_download_button(title = "Download chart data", data = figure_9_data, filename = "figure_9")
```

##### Table

###### Table 9: In England in 2025/26, tirzepatide was the chemical substance with the highest cost

```{r top_10_costs_table}

knitr::kable(table_9, align = "llr")
get_download_button(title = "Download table data", data = figure_9_data, filename = "table_9")
```

</div>

Source: [Additional Tables - Prescription Cost Analysis - A2](`r config$add_anl_link`)

In England in 2025/26:

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "<b>Tirzepatide had the highest cost of £574 million. Tirzepatide is a medication primarily used to treat type 2 diabetes. It is also used to support weight loss in adults with obesity.</b>", width = "100%")` 
::: 
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "<b>Four of the top 10 chemical substances by cost are primarily used as diabetes and weight‑loss drugs.</b>", width = "100%")` 
:::
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "<b>Three of the top 10 chemical substances by cost are primarily used to treat respiratory conditions.</b>", width = "100%")` 
:::
::::

Between 2024/25 and 2025/26, the number of tirzepatide items dispensed rose from 1.1 million to 3.1 million. Over the same period, the average cost per item increased from £108 to £187, adding to the overall increase in costs.

#### `r paste0("The 10 most dispensed chemical substances by total items in ", paste0(substr(max_data_fy,1,5), substr(max_data_fy, 8, 9)))` {- .tabset}

<div class = "tabset">

##### Chart

###### Figure 10: In England in 2025/26, Atorvastatin was the chemical substance with the highest number of items

```{r top_10_items}

figure_10
get_download_button(title = "Download chart data", data = figure_10_data, filename = "figure_10")
```

##### Table

###### Table 10: In England in 2024/25, Atorvastatin was the chemical substance with the highest number of items

```{r top_10_items_table}

knitr::kable(table_10, align = "llr")
get_download_button(title = "Download table data", data = figure_10_data, filename = "table_10")
```

</div>

Source: [Additional Tables - Prescription Cost Analysis - A3](`r config$add_anl_link`)

In England in 2025/26:

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "<b>Atorvastatin was the most dispensed chemical substance with 78 million items. Atorvastatin is a type of statin commonly used to lower cholesterol and help prevent heart disease.</b>", width = "100%")` 
:::
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "<b>Atorvastatin has been the most dispensed chemical substance by items since 2016/17.</b>", width = "100%")` 
:::
:::{.col-md-4} 
`r infoBox_no_border(header = "", text = "<b>Four of the top 10 chemical substances by items belong to BNF Chapter 2 Cardiovascular System. These substances are Atorvastatin, Amlodipine, Ramipril, and Bisoprolol fumarate.</b>", width = "100%")` 
:::
::::

### 2.5. Cost and number of prescription items dispensed by Integrated Care Board (ICB) in England

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-12} 
`r infoBox_border(header = "", text = paste("ICBs are statutory NHS organisations responsible for developing a plan for meeting the health needs of the population, managing the NHS budget, and arranging for the provision of health services in a geographical area. <br><br>It is important to note that the ICB in these statistics is the location of the dispensing contract as listed in NHSBSA administrative records, rather than the prescribing organisation of an item.<br><br>Prescribing and dispensing trends vary regionally, and there are many factors that influence these trends, such as the presence of specialist dispensing contractors within a given area.<br><br>To find out more about ICBs and how they were used in these statistics, please refer to section 3.2 of the [Background Information and Methodology note](",config$background_link,") that accompanies this release.<br><br>For this publication the ICB structure as at the end of March 2026 was used."), width = "100%")` 
::: 
::::

#### `r paste0("Cost per person (£) by ICB ",year_tidy)` {- .tabset}

<div class = "tabset">

##### Chart

###### Figure 11: The mean cost of items dispensed per person in England was £199 in 2025/26

```{r nic_per_pop}
figure_11
get_download_button(title = "Download chart data", data = figure_11_data, filename = "figure_11")
```

##### Table

###### Table 11: The mean cost of items dispensed per person in England was £199 in 2025/26

```{r nic_per_pop_table}
knitr::kable(table_11, align = "lrrr")
get_download_button(title = "Download table data", data = figure_11_data, filename = "table_11")
```

</div>

Source: [Summary Tables - Prescription Cost Analysis - ICB National](`r config$fy_icb_link`) | [Office for National Statistics - Health geographies population estimates](https://www.ons.gov.uk/file?uri=/peoplepopulationandcommunity/populationandmigration/populationestimates/datasets/clinicalcommissioninggroupmidyearpopulationestimates/mid2011tomid2022integratedcareboards2024geography/sapeicb202420112022.xlsx)

In England in 2025/26:

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-6} 
`r infoBox_no_border(header = "", text = "<b>NHS Cambridgeshire and Peterborough ICB continues to have the highest cost per person, this was £420</b>", width = "100%")` 
:::
:::{.col-md-6} 
`r infoBox_no_border(header = "", text = "<b>NHS South West London Integrated Care Board has the lowest cost per person, this was £137.</b>", width = "100%")` 
:::
::::

As in 2024/25, NHS Cambridgeshire and Peterborough ICB's cost per person is more than double the national average. Dispensing contractors in this ICB offer specialist stoma care, 35% of their total costs of £401 million in 2025/26 were in BNF Chapter 23 Stoma Appliances.

#### `r paste0("Dispensed items per person by ICB ",year_tidy)` {- .tabset}

<div class = "tabset">

##### Chart

###### Figure 12: The mean number of items dispensed per person in England was 22 in 2025/26

```{r items_per_pop}

figure_12
get_download_button(title = "Download chart data", data = figure_12_data, filename = "figure_12")  
```

##### Table

###### Table 12: The mean number of items dispensed per person in England was 22 in 2025/26

```{r items_per_pop_table}
knitr::kable(table_12, align = "lrrr")
get_download_button(title = "Download table data", data = figure_12_data, filename = "table_12")
```

</div>

Source: [Summary Tables - Prescription Cost Analysis - ICB National](`r config$fy_icb_link`) | [Office for National Statistics - Mid-year population estimates](https://www.ons.gov.uk/file?uri=/peoplepopulationandcommunity/populationandmigration/populationestimates/datasets/clinicalcommissioninggroupmidyearpopulationestimates/mid2011tomid2022integratedcareboards2024geography/sapeicb202420112022.xlsx)

In England in 2025/26:

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-6} 
`r infoBox_no_border(header = "", text = "<b>NHS West Yorkshire ICB continues to have the highest dispensed items per person, this was 39.</b>", width = "100%")` 
:::
:::{.col-md-6} 
`r infoBox_no_border(header = "", text = "<b>NHS Surrey Heartlands ICB continues to have the lowest dispensed items per person, this was 14.</b>", width = "100%")` 
:::
::::

### 2.6. Top presentations by changes in unit and total costs

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-12} 
`r infoBox_border(header = "", text = "Presentation is the name given to the specific type, strength, and formulation of a drug, or the specific type of an appliance. For example, paracetamol 500mg tablets.<br><br>The charts in this section have been limited to presentations with a total cost in 2024/25 greater than £1 million.<br><br>A unit cost for a presentation is calculated by dividing the total cost of items dispensed in the period by the total quantity.<br><br>Price concessions can be granted by the Department of Health and Social Care (DHSC) for medicines that pharmacists are struggling to obtain at or below the price set out in the Drug Tariff. These concessions can have large impacts on the unit costs of drugs.", width = "100%")` 
::: 
::::

#### `r paste0("Top 10 presentations by percentage increase in unit cost ",paste0(substr(max_data_fy_minus_1,1,5), substr(max_data_fy_minus_1, 8, 9)), " to ", paste0(substr(max_data_fy,1,5), substr(max_data_fy, 8, 9)))` {- .tabset}

<div class = "tabset">

##### Chart

###### Figure 13: Paxlovid 150mg/100mg tablets had the highest increase in unit cost from 2024/25 to 2025/26

```{r unit_cost_increase}
figure_13
get_download_button(title = "Download chart data", data = figure_13_data, filename = "figure_13")
```

##### Table

###### Table 13: Paxlovid 150mg/100mg tablets had the highest increase in unit cost from 2024/25 to 2025/26

```{r unit_cost_increase_table}
knitr::kable(table_13, align = "llr")
get_download_button(title = "Download table data", data = figure_13_data, filename = "table_13")
```

</div>

Source: [Additional Tables - Prescription Cost Analysis - A11](`r config$add_anl_link`)

Paxlovid 150mg/100mg tablets had the highest increase in unit cost from 2024/25 to 2025/26, an increase of 30,768%. 

Between April 2025 and May 2025 the unit cost of Paxlovid 150mg/100mg increased from around 8 pence per tablet to around £28 per tablet. This is as the manufacturer set a new list price per pack of £829.

#### `r paste0("Top 10 presentations by percentage decrease in unit cost ",paste0(substr(max_data_fy_minus_1,1,5), substr(max_data_fy_minus_1, 8, 9)), " to ", paste0(substr(max_data_fy,1,5), substr(max_data_fy, 8, 9)))`  {- .tabset}

<div class = "tabset">

##### Chart

###### Figure 14: Rivaroxaban 20mg tablets had the highest decrease in unit cost from 2024/25 to 2025/26

```{r unit_cost_decrease}
figure_14
get_download_button(title = "Download chart data", data = figure_14_data, filename = "figure_14")
```

##### Table

###### Table 14: Rivaroxaban 20mg tablets had the highest decrease in unit cost from 2024/25 to 2025/26

```{r unit_cost_decrease_table}
knitr::kable(table_14, align = "llr")
get_download_button(title = "Download table data", data = figure_14_data, filename = "table_14")
```

</div>

Source: [Additional Tables - Prescription Cost Analysis - A12](`r config$add_anl_link`)

Rivaroxaban 20mg tablets had the highest decrease in unit cost from 2024/25 to 2025/26, a decrease of 93%. Rivaroxaban came off patent and was made a generic in 2024.

#### `r paste0("Top 10 presentations by increase in total cost ", paste0(substr(max_data_fy_minus_1,1,5), substr(max_data_fy_minus_1, 8, 9)), " to ", paste0(substr(max_data_fy,1,5), substr(max_data_fy, 8, 9)))` {- .tabset}

<div class = "tabset">

##### Chart

###### Figure 15: FreeStyle Libre 2 Plus Sensor kit had the highest increase in total costs from 2024/25 to 2025/26

```{r abs_cost_increase}
figure_15
get_download_button(title = "Download chart data", data = figure_15_data, filename = "figure_15")
```

##### Table

###### Table 15: FreeStyle Libre 2 Plus Sensor kit had the highest increase in total costs from 2024/25 to 2025/26

```{r abs_cost_increase_table}
knitr::kable(table_15, align = "llr")
get_download_button(title = "Download table data", data = figure_15_data, filename = "table_15")
```

</div>

Source: [Additional Tables - Prescription Cost Analysis - A13](`r config$add_anl_link`)

Between 2024/25 and 2025/26, the total costs of FreeStyle Libre 2 Plus Sensor kits increased by £255 million, the highest increase of all presentations. FreeStyle Libre 2 Plus Sensor kits continuous glucose monitoring sensors are used by people with diabetes to track glucose levels without routine finger‑prick testing. 

Dapagliflozin 10mg tablets had the second largest increase in total costs between 2024/25 and 2025/26. Dapagliflozin is the generic form of Forxiga which came off patent in 2025 and is shown in figure 16 to have the highest decrease in total costs over the same period.

#### `r paste0("Top 10 presentations by decrease in total cost ", paste0(substr(max_data_fy_minus_1,1,5), substr(max_data_fy_minus_1, 8, 9)), " to ", paste0(substr(max_data_fy,1,5), substr(max_data_fy, 8, 9)))` {- .tabset}

<div class = "tabset">

##### Chart

###### Figure 16: Forxiga 10mg tablets had the highest decrease in total costs from 2024/25 to 2025/26

```{r abs_cost_decrease}
figure_16
get_download_button(title = "Download chart data", data = figure_16_data, filename = "figure_16")
```

##### Table

###### Table 16: Forxiga 10mg tablets had the highest decrease in total costs from 2024/25 to 2025/26

```{r abs_cost_decrease_table}
knitr::kable(table_16, align = "llr")
get_download_button(title = "Download table data", data = figure_16_data, filename = "table_16")
```

</div>

Source: [Additional Tables - Prescription Cost Analysis - A14](`r config$add_anl_link`)

Between 2024/25 and 2025/26, the total costs of Forxiga 10mg tablets decreased by £249 million, the highest decrease of all presentations. Forxiga came off patent and was made a generic in 2025.

---

## 3. UK comparisons

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-12} 
`r infoBox_border(header = "", text = paste("Prescription Cost Analysis statistics are produced separately by each UK nation, and there are differences in data collection, healthcare structures, and methodology. Therefore, it's important to be cautious when comparing PCA data of different UK nations, and to consider the broader context of the UK's health and social care landscape.<br><br>We have chosen a few high-level measures that allow for broad comparisons across the devolved administrations. For more detailed information, please refer to each country's official PCA publication<br><br>Our comparisons are based on the most recent PCA data available at the time of analysis and use the Office for National Statistics (ONS) 2024 mid-year population estimates to calculate per-person figures for the 2024/25 financial year.<br><br>For Wales, the figures reflect prescriptions issued by Welsh GPs and dispensed either in Wales or elsewhere in the UK.<br><br>For Northern Ireland the PCA data used is for the calendar year 2024."), width = "100%")` 
::: 
::::

#### `r paste0("Cost per person for the UK nations ", max_data_fy_minus_1)` {- .tabset}

<div class = "tabset">

##### Chart

###### Figure 17: England had the lowest prescription item cost per person in 2024/25

```{r uk_comparisons_costs}
figure_17
get_download_button(title = "Download chart data", data = figure_17_data, filename = "figure_17")

```

##### Table

###### Table 17: England had the lowest prescription item cost per person in 2024/25

```{r uk_comparisons_costs_table}
knitr::kable(table_17, align = "lr")
get_download_button(title = "Download table data", data = figure_17_data, filename = "table_17")
```

</div>

Source: [Additional Tables - Prescription Cost Analysis - A1](`r config$add_anl_link`) | [Public Health Scotland - Dispenser payments and prescription cost analysis](`r config$scotland_pca`) | [Welsh Government - Prescriptions in Wales](`r config$wa_pca_link`) | [HSC Business Services Organisation - General pharmaceutical services and prescribing statistics](`r config$ni_pca_link`) | [Office for National Statistics - Mid-year population estimates](https://www.ons.gov.uk/peoplepopulationandcommunity/populationandmigration/populationestimates/datasets/populationestimatestimeseriesdataset)

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-6} 
`r infoBox_no_border(header = "", text = "<b>England had the lowest prescription item cost per person in 2024/25 at £190.</b>", width = "100%")` 
::: 
:::{.col-md-6} 
`r infoBox_no_border(header = "", text = "<b>Northern Ireland had the highest prescription item cost per person in 2024/25 at £261.</b>", width = "100%")` 
::: 
::::

#### `r paste0("Items per person for the UK nations ", max_data_fy_minus_1)` {- .tabset}

<div class = "tabset">

##### Chart

###### Figure 18: Scotland had the fewest items dispensed per person in 2024/25

```{r uk_comparisons_items}

figure_18
get_download_button(title = "Download chart data", data = figure_18_data, filename = "figure_18")
```

##### Table

###### Table 18: Scotland had the fewest items dispensed per person in 2024/25

```{r uk_comparisons_items_table}
knitr::kable(table_18, align = "lr")
get_download_button(title = "Download table data", data = figure_18_data, filename = "table_18")
```

</div>

Source: [Additional Tables - Prescription Cost Analysis - A1](`r config$add_anl_link`) | [Public Health Scotland - Dispenser payments and prescription cost analysis](`r config$scotland_pca`) | [Welsh Government - Prescriptions in Wales](`r config$wa_pca_link`) | [HSC Business Services Organisation - General pharmaceutical services and prescribing statistics](`r config$ni_pca_link`) | [Office for National Statistics - Mid-year population estimates](https://www.ons.gov.uk/peoplepopulationandcommunity/populationandmigration/populationestimates/datasets/populationestimatestimeseriesdataset)

::::{.row style="display: flex; margin-bottom: 24px;"} 
:::{.col-md-6} 
`r infoBox_no_border(header = "", text = "<b>Scotland had the lowest number of prescription items dispensed per person in 2024/25 at 21 items.</b>", width = "100%")` 
::: 
:::{.col-md-6} 
`r infoBox_no_border(header = "", text = "<b>Wales had the highest number of prescription items dispensed per person in 2024/25 at 27 items.</b>", width = "100%")` 
::: 
::::


---

## 4. About these statistics

### 4.1. Monthly PCA data {.toc-ignore}

The NHSBSA publishes a [monthly administrative management information dataset for PCA](`r config$nhsbsa_pca_link`) using the same methodology as these statistics. However, while our annual PCA is an Official Statistic, the monthly dataset is not.

There is a slight difference between the annual National Statistics and the monthly administrative data for PCA. The annual publication uses the geographical structure as of the 31 March for that year, while the monthly data uses the structure at the end of the month. If a dispensing contractor changes parent organisation during the year, this may affect the ICB level total in the aggregated monthly data compared to the annual National Statistics, but it will not affect the national totals. However, such cases are rare.


### 4.2. Uses of these statistics {.toc-ignore}

These statistics serve several purposes, including:

* providing a national overview of community prescription costs and volumes in England for both the financial year and the calendar year

* analysing cost and volume trends across different levels of the BNF hierarchy

* allowing public scrutiny of national prescribing habits

* monitoring uptake of new to market proprietary presentations

* monitoring the proportion of prescriptions dispensed generically

### 4.3. Changes to this publication {.toc-ignore}

Information on items dispensed via the [PCS](https://www.england.nhs.uk/primary-care/pharmacy/pharmacy-services/nhs-pharmacy-contraception-service/) and the [Seasonal Influenza Vaccination Service](https://www.england.nhs.uk/wp-content/uploads/2025/07/prn01998-v-community-pharmacy-advanced-service-specification-seasonal-influenza-vaccination-1-sept-2025-31-march-2026.pdf) is included in the BNF presentation level data and SNOMED level data tables of the supplementary tables.

For full details of the changes made please refer to the [Background Information and Methodology note](`r config$background_link`) released alongside these statistics.

### 4.4. Timeliness of this publication {.toc-ignore}

This publication is released in June of each year, when March data becomes available. This allows us to provide a full financial year overview. Calendar year data is published in February of each year when December data becomes available on the [NHSBSA Open Data Portal](https://opendata.nhsbsa.net/) (ODP).

### 4.5. Price concessions {.toc-ignore}

When community pharmacies cannot source a drug at or below the reimbursement price as set out in the Drug Tariff, the DHSC can introduce a [price concession](https://cpe.org.uk/funding-and-reimbursement/reimbursement/price-concessions/) at the request of Community Pharmacy England. A price concession can be requested for any drugs listed in Part VIIIA, Part VIIIB, and Part VIIID of the Drug Tariff. For any drugs granted price concessions, contractors are automatically reimbursed at the new prices for that month.

---

## 5. Rounding

The high-level figures in this statistical summary have been rounded as per the table below:

```{r rounding_table}
rounding_table <- data.table(
  "From" = c("0", "1,001", "10,001", "100,001", "1,000,001", "10,000,001", "100,000,001", "10,000,000,001"),
  "To" = c("1,000", "10,000", "100,000", "1,000,000", "10,000,000", "100,000,000", "10,000,000,000", "100,000,000,000"),
  "Round to nearest" = c("1", "100", "1,000", "10,000", "100,000", "1,000,000", "10,000,000", "100,000,000")
)

knitr::kable(rounding_table, align = "rrr")
```

All changes and totals are calculated prior to rounding. Percentage changes are calculated prior to rounding and then are rounded to the nearest whole number. Since all figures within this statistical summary have been rounded, they may not match totals elsewhere when aggregated.

The summary tables released with this publication allow users to investigate this data at lower levels of granularity. Figures in the supplementary tables have not been rounded.

---

## 6. Accessibility

If you need information on this website in a different format, such as accessible PDF, large print, easy read, audio recording or braille, you can contact us by:

**Email**: accessibility@nhsbsa.nhs.uk

**Telephone**: 0191 203 5318

[Find out about call charges](https://www.nhsbsa.nhs.uk/contact-us/call-charges-and-phone-numbers)

We will consider your request and get back to you in 5 working days.

These contact details are only for accessibility queries. This email address is not for technical queries or IT problems. If you have a query that is not about accessibility, go to the ‘Feedback and contact us’ section of this page.

Read our [Accessibility statement for Official Statistics Narratives](https://www.nhsbsa.nhs.uk/accessibility-statement-official-statistics-narratives).

---

## 7. Feedback and contact us

Feedback is important to us. We welcome any questions and comments relating to these statistics. 

You can complete a [short survey about this publication](`r config$stats_survey_link`) to help us improve the Official Statistics that we produce. All responses will remain anonymous, and individuals will not be identifiable in any report that we produce.

You can view our [privacy policy](https://www.nhsbsa.nhs.uk/our-policies/privacy) on our website to see how your data is used and stored.

You can contact us by:

**Email:** statistics@nhsbsa.nhs.uk

**You can also write to us at:**

NHSBSA - Statistics  
NHS Business Services Authority  
Stella House  
Goldcrest Way  
Newburn Riverside  
Newcastle upon Tyne  
NE15 8NY

**Responsible statistician:** `r config$responsible_statistician`

</main>